### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd in data section

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Data Overview

This analysis uses historical stock price data for FAANG companies (Meta, Amazon, Apple, Netflix, and Google) obtained from the Yahoo Finance API via the `tidyquant` library in R on November 30, 2024. The dataset spans from January 1, 2020, to December 31, 2024, providing daily records of stock performance, including key metrics such as opening, high, low, and closing prices, trading volume, and adjusted closing prices normalized for corporate actions. To enhance the predictive power of the data, several derived features were constructed. These include lagged adjusted prices to capture sequential trends, daily returns to measure proportional changes, rolling averages to smooth short- and long-term fluctuations, and a rolling standard deviation of prices to quantify market volatility. These features offer a structured framework for identifying patterns and forecasting future stock movements.

The raw data underwent preprocessing to address missing values, particularly for adjusted prices. Missing entries were filled using forward and backward imputation to ensure data continuity. Placeholder dates for December 2024 were interpolated to prepare for out-of-sample predictions. The cleaned dataset, prepared using the `tidyverse`, `lubridate`, `TTR`, and `arrow` libraries, provides a consistent and comprehensive foundation for modeling and analysis.

Further details on the data preparation process are included in [Appendix -@sec-data-cleaning].

## Measurement
Stock price data encapsulates market dynamics influenced by diverse factors, including macroeconomic conditions, corporate events, and investor sentiment. The raw variables in this dataset, such as **`open`**, **`high`**, **`low`**, **`close`**, and **`adjusted`**, capture key price points and provide a snapshot of market activity for each trading day. However, while these variables accurately reflect daily trading outcomes, they may not fully represent intraday fluctuations or complex trading dynamics.

Constructed variables enhance the analytical depth of the dataset by quantifying sequential trends, price variability, and momentum. For instance, **`Lag_1`** captures the adjusted closing price from the previous trading day, providing context for sequential patterns. Variables like **`daily_return`** measure proportional daily changes in price, while moving averages such as **`Rolling_Mean_7`**, **`sma_20`**, and **`sma_50`** smooth short- and long-term price fluctuations, offering a clearer view of overall trends. Additionally, **`volatility`**, computed as the rolling standard deviation of closing prices, quantifies recent price variability and reflects market uncertainty.

Market activity is further contextualized through **`volume`**, representing the total shares traded daily, a proxy for investor interest and liquidity. The inclusion of **`symbol_encoded`**, a numerical encoding of stock tickers, standardizes categorical data for compatibility with machine learning models. Together, these features create a robust framework for analysis, balancing raw price data with derived metrics that highlight market trends and dynamics.

Despite the strengths of these variables, limitations exist. Raw price points do not account for intraday variations, while constructed metrics may oversimplify complex market behaviors. External shocks, such as economic crises or regulatory changes, may not be fully captured within the variables’ scope. Acknowledging these constraints is essential for interpreting the results and understanding the broader context of the analysis. These limitations are further discussed in {@sec-data-details}.

## Variables

The dataset used for this analysis includes both original and constructed variables, designed to capture temporal trends, volatility, and momentum in stock price movements. Below is a detailed overview of the variables included in the model:

### Original Variables
The original variables directly obtained from the raw stock price data are:

- **`symbol`**: The stock ticker symbol identifying the company (e.g., AAPL for Apple, AMZN for Amazon).
- **`date`**: The trading date, essential for analyzing time-series trends.
- **`open`**: The stock’s opening price for the trading day, indicating initial market conditions.
- **`high`**: The highest price reached during the trading day, reflecting intraday volatility.
- **`low`**: The lowest price during the trading day, reflecting downward market trends.
- **`close`**: The stock’s unadjusted closing price, representing the final value at the end of trading.
- **`volume`**: The total number of shares traded during the day, reflecting market activity.
- **`adjusted`**: The closing price adjusted for corporate actions such as splits or dividends, ensuring consistency for trend analysis.

### Constructed Variables
The constructed variables were engineered to enhance the model's predictive performance by capturing price trends, volatility, and momentum:

- **`Lag_1`**: The closing price from the previous trading day, derived using the `lag()` function. This variable captures sequential patterns in stock prices.
- **`Rolling_Mean_7`**: A 7-day moving average of adjusted closing prices, computed using the `rollmean()` function. This variable smooths short-term fluctuations to highlight weekly trends.
- **`sma_20`**: The 20-day simple moving average (SMA) of adjusted prices, reflecting medium-term price trends.
- **`sma_50`**: The 50-day simple moving average of adjusted prices, providing insights into longer-term trends.
- **`volatility`**: The 20-day rolling standard deviation of adjusted prices, capturing short-term price variability.
- **`daily_return`**: The percentage change in adjusted closing prices relative to the previous day, calculated as:
  $$
  \text{daily_return} = \frac{\text{adjusted}}{\text{Lag_1}} - 1
  $$
  This variable normalizes price changes for comparability across time and stocks.
- **`symbol_encoded`**: A numeric encoding of the stock ticker symbol, allowing the categorical `symbol` variable to be included in the modeling process.

### Outcome Variable
The **outcome variable** for this analysis is the **adjusted closing price (`adjusted`)**, which serves as the target for prediction. This represents the stock's closing price adjusted for corporate actions like stock splits or dividends, providing a normalized measure of stock value. The model aims to predict this variable based on historical data and constructed predictors.

### Predictor Variables
The predictor variables used to model the adjusted closing price are:

- **`Lag_1`**: The prior day's adjusted closing price, reflecting momentum.
- **`Rolling_Mean_7`**: A smoothed metric capturing weekly price trends.
- **`sma_20`** and **`sma_50`**: Simple moving averages representing medium- and long-term price trends.
- **`volatility`**: The short-term price variability, highlighting potential risks.
- **`daily_return`**: The normalized daily percentage price change.
- **`symbol_encoded`**: A numerical representation of the stock ticker symbol, capturing company-level differences.

Together, these variables provide a robust framework for understanding and predicting stock price movements over time, leveraging both raw market indicators and derived metrics.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-12-03 paper/paper.qmd
committing changes to results section in paper.qmd

@@
 ---
-title: "My title"
+title: "Predicting FAANG Stock Prices: A Data-Driven Approach Using Machine Learning Models"
 subtitle: "My subtitle if needed"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Jimin (Jamie) Lee
+thanks: "Code and data are available at: https://github.com/jamiejiminlee/FAANG-Stock-Forecast.git."
 date: today
@@
 #| message: false
+#| include: false
+#| warning: false
+#| message: false
 
 library(tidyverse)
-library(palmerpenguins)
+library(ggplot2)
+library(arrow)
+library(kableExtra)
+library(xgboost)
+library(tidyr)
+library(zoo)
+library(scales)
+library(reshape2)
+
+analysis_data <- read_parquet("/Users/jamielee/TechStockForecast/data/02-analysis_data/analysis_data.parquet")
 ```
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+library(xgboost)
+
+nov_model <- readRDS("/Users/jamielee/TechStockForecast/models/nov_model.rds")
+dec_model <- readRDS("/Users/jamielee/TechStockForecast/models/dec_model.rds")
+
+```
+
 ### Model justification
@@
 # Results
+```{r}
+# Filter November data for prediction
+nov_test <- analysis_data %>%
+  filter(as.Date(date) >= as.Date("2024-11-01") & as.Date(date) <= as.Date("2024-11-30"))
 
-Our results are summarized in @tbl-modelresults.
+# Prepare testing features for November
+X_test_nov <- nov_test %>%
+  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+  as.matrix()
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+# Predict November prices
+dtest_nov <- xgb.DMatrix(data = X_test_nov)
+nov_predictions <- predict(nov_model, dtest_nov)
 
-library(rstanarm)
+# Combine predictions with actual November data
+nov_results <- nov_test %>%
+  mutate(predicted_adjusted = nov_predictions) %>%
+  select(date, symbol, adjusted, predicted_adjusted)
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+# Filter December data for prediction
+dec_test <- analysis_data %>%
+  filter(as.Date(date) >= as.Date("2024-12-01") & as.Date(date) <= as.Date("2024-12-31"))
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
+# Prepare testing features for December
+X_test_dec <- dec_test %>%
+  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+  as.matrix()
+
+# Predict December prices
+dtest_dec <- xgb.DMatrix(data = X_test_dec)
+dec_predictions <- predict(dec_model, dtest_dec)
+
+# Combine predictions with actual December data
+dec_results <- dec_test %>%
+  mutate(predicted_adjusted = dec_predictions) %>%
+  select(date, symbol, predicted_adjusted)
+
+```
+
+## Actual vs Predicted for November 2024
+```{r}
+
+# Plot actual vs predicted prices for November 2024
+ggplot(nov_results, aes(x = as.Date(date))) +
+  geom_line(aes(y = adjusted, color = "Actual"), size = 1) +  # Actual prices
+  geom_line(aes(y = predicted_adjusted, color = "Predicted"), linetype = "dashed", size = 1) +  # Predicted prices
+  facet_wrap(~symbol, scales = "free_y") +  # Separate panels for each stock
+  labs(
+    title = "Actual vs Predicted Stock Prices for November 2024",
+    x = "Date",
+    y = "Stock Price",
+    color = "Legend"
+  ) +
+  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) +  # Define colors
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    text = element_text(size = 12),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+```
+
+## Heatmap of Prediction Errors (November 2024)
+```{r}
+# Add the `error` column to `nov_results`
+nov_results <- nov_results %>%
+  mutate(error = abs(adjusted - predicted_adjusted))  # Calculate absolute error
+
+# Prepare data for heatmap
+nov_heatmap_data <- nov_results %>%
+  select(date, symbol, error) %>%
+  spread(symbol, error)
+
+# Plot heatmap of prediction errors
+ggplot(melt(nov_heatmap_data, id.vars = "date"), aes(x = date, y = variable, fill = value)) +
+  geom_tile() +
+  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
+  labs(
+    title = "Heatmap of Prediction Errors for November 2024",
+    x = "Date",
+    y = "Stock Symbol",
+    fill = "Error"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    text = element_text(size = 12)
+  )
+```
+
+## Average Predicted Stock Price - November
+```{r}
+# Calculate average predicted prices for November
+nov_avg_prices <- nov_results %>%
+  group_by(symbol) %>%
+  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE))
+
+# Plot average predicted prices
+ggplot(nov_avg_prices, aes(x = symbol, y = avg_predicted_price, fill = symbol)) +
+  geom_bar(stat = "identity", width = 0.6) +
+  labs(
+    title = "Average Predicted Stock Prices for November 2024",
+    x = "Stock Symbol",
+    y = "Average Predicted Price",
+    fill = "Stock Symbol"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "none",
+    text = element_text(size = 12),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  ) +
+  scale_fill_brewer(palette = "Set2")
+
+```
+
+## Predicted Stock Price Trends - December
+```{r}
+ggplot(dec_results, aes(x = as.Date(date), y = predicted_adjusted, color = symbol)) +
+  geom_line(size = 1) +
+  labs(
+    title = "Predicted Stock Price Trends for December 2024",
+    x = "Date",
+    y = "Predicted Stock Price",
+    color = "Stock Symbol"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    text = element_text(size = 12),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+```
+
+## Average Stock Prices Bar Chart - December
+```{r}
+# Calculate average predicted prices for December
+dec_avg_prices <- dec_results %>%
+  group_by(symbol) %>%
+  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE))
+
+# Plot average predicted prices
+ggplot(dec_avg_prices, aes(x = symbol, y = avg_predicted_price, fill = symbol)) +
+  geom_bar(stat = "identity", width = 0.6) +
+  labs(
+    title = "Average Predicted Stock Prices for December 2024",
+    x = "Stock Symbol",
+    y = "Average Predicted Price",
+    fill = "Stock Symbol"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "none",
+    text = element_text(size = 12),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  ) +
+  scale_fill_brewer(palette = "Set2")
+
+```
+
+## Average Predicted Stock Price - December
+
+```{r}
+# Calculate average predicted price for December
+dec_avg_prices <- dec_results %>%
+  group_by(symbol) %>%
+  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE)) %>%
+  mutate(avg_predicted_price = dollar(round(avg_predicted_price, 2))) %>%  # Format as dollar and round
+  arrange(desc(avg_predicted_price))  # Sort by average price in descending order
+
+# Display the table
+dec_avg_prices %>%
+  kable(
+    caption = "Average Predicted Stock Prices for December 2024",
+    col.names = c("Stock Symbol", "Average Predicted Price"),
+    format = "markdown"  # Output as markdown for readability
+  )
+```
+
+## Combined November and December
+```{r}
+# Combine November and December results
+combined_results <- bind_rows(
+  nov_results %>% mutate(month = "November"),
+  dec_results %>% mutate(month = "December")
+)
+
+# Plot predicted prices with facets for each stock
+ggplot(combined_results, aes(x = as.Date(date), y = predicted_adjusted, color = month)) +
+  geom_line(size = 1) +
+  facet_wrap(~symbol, scales = "free_y") +
+  labs(
+    title = "Predicted Stock Prices for November and December 2024",
+    x = "Date",
+    y = "Predicted Stock Price",
+    color = "Month"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    text = element_text(size = 12),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+
+```
+
+## Average Predicted Stock Prices for November and December 2024
+
+```{r}
+# Reorder the month factor so that "November" comes before "December"
+avg_predicted_prices <- avg_predicted_prices %>%
+  mutate(month = factor(month, levels = c("November", "December")))
+
+# Plot with the updated order
+ggplot(avg_predicted_prices, aes(x = symbol, y = mean_price, fill = month)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  scale_fill_manual(values = c("November" = "#F8766D", "December" = "#00BFC4")) +
+  labs(
+    title = "Average Predicted Stock Prices for November and December 2024",
+    x = "Stock Symbol",
+    y = "Average Predicted Price (in USD)",
+    fill = "Month"
   )

### 2024-12-03 paper/paper.qmd
committing changes to result section of paper.qmd

@@
 ```{r}
+# Calculate average predicted prices for November and December
+avg_predicted_prices <- combined_results %>%
+  group_by(symbol, month) %>%
+  summarize(avg_price = mean(predicted_adjusted, na.rm = TRUE), .groups = "drop")
+
 # Reorder the month factor so that "November" comes before "December"
@@
 # Plot with the updated order
-ggplot(avg_predicted_prices, aes(x = symbol, y = mean_price, fill = month)) +
+ggplot(avg_predicted_prices, aes(x = symbol, y = avg_price, fill = month)) +
   geom_bar(stat = "identity", position = "dodge") +

### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd in model section

@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+## Model Overview
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The models in this study forecast the adjusted closing prices of FAANG stocks (Meta, Amazon, Apple, Netflix, and Google) for the months of November and December 2024.
+Separate models are constructed for each month to capture the time-dependent nature of stock price movements and to account for fluctuations in market conditions leading up to the forecast period.
+These models utilize a combination of historical stock price data and technical indicators, such as moving averages, volatility, and lagged price features, to estimate future stock prices for each of the five FAANG companies.
 
-## Model set-up
+To model the changes in stock prices over time, we use features that capture both short-term and long-term market trends.
+The models incorporate lagged stock prices, 7-day rolling averages, 20-day and 50-day simple moving averages (SMA), daily returns, and volatility indicators.
+These features are crucial for understanding the patterns and cycles in stock price movements.
+XGBoost, a powerful machine learning algorithm, is used to build these models, allowing for the capture of non-linear relationships and interactions between the technical indicators and future stock prices.
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+The training datasets for the November and December models are defined using data up to October 31, 2024, and November 30, 2024, respectively.
+The models take into account temporal changes in stock prices by using features like lagged prices and rolling averages, which help capture the dynamics of stock price behavior.
+Additionally, XGBoost's flexibility allows the models to account for varying levels of volatility and market shifts across the different FAANG stocks.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+This modeling approach enables the prediction of future stock prices, considering how past price trends and market indicators evolve.
+By focusing on the November and December timeframes, the models aim to provide forecasts that reflect recent market conditions and predict likely price movements as we approach the end of 2024.
+The detailed model specifications, including feature selection and hyperparameter tuning, are further discussed in [Appendix -@sec-model-details]
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+## Model Assumptions
+
+To ensure the validity and reliability of our stock price prediction models, several key assumptions are made during the modeling process.
+These assumptions are necessary to capture the complexities of stock price behavior while maintaining a clear and interpretable framework for forecasting:
+
+- **Non-Linearity in Price Movements**: Stock prices often exhibit non-linear patterns due to various market forces, including economic news, company performance, and broader market trends.
+To account for these non-linearities, we use XGBoost, which allows for the modeling of complex, non-linear relationships between the predictors (such as lagged prices and moving averages) and the target variable (adjusted closing prices).
+This assumption ensures that the model can effectively capture the changing dynamics of stock price movements over time.
+
+- **Stationarity of Price Features**: The model assumes that the selected technical indicators—such as moving averages, volatility, and daily returns—are stationary over the forecasting period.
+While stock prices are inherently volatile, we assume that the patterns observed in historical data are stable enough to provide reliable predictions for the future months.
+This assumption allows us to model stock prices without explicitly accounting for trends or seasonality, assuming that the past behavior is a good representation of future dynamics.
+
+- **Independence of Observations**: Each daily stock price observation is assumed to be independent of the others.
+While stock prices are influenced by previous values (as captured through lagged features), we assume that each day's stock price data point does not directly depend on other data points in ways not captured by the lagged variables or technical indicators.
+This assumption is important for simplifying the modeling process and ensuring that the features can be treated as independent predictors.
+
+- **Stationarity of Model Parameters**: We assume that the relationships between the stock prices and the chosen technical indicators (e.g., moving averages, volatility) remain relatively stable over the forecast horizon.
+While stock prices can be influenced by many unpredictable factors, this assumption allows the model to generalize well to the forecasting periods (November and December 2024).
+
+- **Feature Importance**: We assume that the selected features—lagged prices, moving averages, volatility, and daily returns—capture the most important aspects of stock price movements for the FAANG companies.
+While other factors (such as macroeconomic events, news sentiment, or geopolitical developments) may influence stock prices, these features are chosen for their relevance to technical analysis and their ability to capture short-term and long-term market trends.
+
+- **Use of XGBoost’s Flexibility**: The XGBoost algorithm is assumed to be an appropriate method for this analysis due to its ability to handle large datasets, capture non-linear relationships, and deal with complex interactions between features.
+We assume that its regularization techniques will help prevent overfitting while still capturing the intricate patterns present in the stock price data.
+
+These assumptions form the foundation of the models and guide the interpretation of the results.
+While they allow for effective forecasting, it is important to acknowledge that stock price prediction is inherently uncertain, and the accuracy of predictions may vary depending on unforeseen market events and fluctuations.
+
+## Model Setup
+
+Modeling processes are conducted by employing the following packages: `tidyquant` package of @tidyquant for downloading and managing financial data, `xgboost` package of @xgboost for implementing the machine learning model, `dplyr` package of @dplyr for data manipulation, and `lubridate` package of @lubridate for handling dates and times.
+
+### November Model
+
+The primary goal of the November model is to predict the adjusted closing prices of FAANG stocks for the month of November 2024 based on historical stock price data.
+This model is structured as a regression task using the XGBoost algorithm to estimate the future adjusted closing prices for each of the five FAANG companies.
+
+The model setup can be described as follows:
+
+$$
+\hat{y}_{i} = f(X_i)
+$$
+
+Where:
+
+- $\hat{y}_i$ is the predicted adjusted closing price for the stock on day $i$.
+- $X_i$ represents the feature set for day $i$, which includes:
+  - `Lag_1`: The adjusted closing price of the previous day.
+  - `Rolling_Mean_7`: The 7-day rolling average of the adjusted closing price.
+  - `sma_20`: The 20-day simple moving average.
+  - `sma_50`: The 50-day simple moving average.
+  - `volatility`: The 14-day rolling standard deviation of daily returns.
+  - `daily_return`: The daily percentage change in the adjusted closing price.
+  - `symbol_encoded`: A numeric encoding for the FAANG companies (Meta, Amazon, Apple, Netflix, Google).
+
+The dataset for the November model is prepared by extracting the relevant data from January 2020 to October 31, 2024.
+Features like moving averages and volatility are calculated based on this historical data, and the model is trained to predict the adjusted closing prices for November 2024.
+
+The model was implemented using the **`xgboost`** package in R, which is optimized for performance in regression tasks:
+
+$$
+y_i = \text{XGBoost}(X_i, \theta)
+$$
+
+Where $y_i$ represents the actual adjusted closing price for day $i$ and $\theta$ represents the model parameters to be learned during training. The model is trained with the following key parameters:
+- **Max Depth**: 6
+- **Learning Rate**: 0.1
+- **Number of Rounds**: 100
+- **Objective Function**: Regression with squared error
+
+### December Model
+
+The December model follows a similar approach to the November model, but is trained using data up to November 30, 2024.
+This ensures that the model accounts for the most recent stock price movements leading up to December, providing a more accurate prediction for the final month of 2024.
+
+The setup for the December model mirrors that of the November model, with the same features and target variable, but applied to the updated dataset.
+The model structure remains the same, and the XGBoost algorithm is again used for regression:
+
+$$
+\hat{y}_{i} = f(X_i)
+$$
+
+Where:
+
+- $\hat{y}_i$ is the predicted adjusted closing price for the stock on day $i$ in December 2024.
+- $X_i$ includes the same features as the November model, but with data up to November 30, 2024.
+
+The December model is trained using the same XGBoost parameters as the November model.
+Once trained, the model will generate predictions for December 2024, leveraging the latest market conditions reflected in the training data.
+
+The assumptions and model design are consistent with the November model, with the key differences being the updated data and the forecast period.
+By training separate models for November and December, we can capture any evolving trends or shifts in stock price behavior as we approach the end of 2024.
 
@@
 
-### Model justification
+## Model Justification
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The XGBoost model is the most suitable approach for forecasting FAANG stock prices in November and December 2024.
+Our primary outcome variable is the adjusted closing price of each stock, which is continuous and unbounded.
+XGBoost, a gradient boosting algorithm, excels in handling regression tasks with large datasets, capturing complex, non-linear relationships between features such as lagged prices, moving averages, volatility, and daily returns.
+To model these relationships effectively, XGBoost's ability to handle feature interactions and non-linearities is key.
+The inclusion of rolling averages and volatility allows the model to account for both short-term fluctuations and longer-term trends in stock prices, which are often influenced by broader market dynamics.
+Additionally, the model's flexibility helps capture the intricacies of stock price behavior over time, ensuring that past market patterns are reflected in future predictions.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+While simpler models like linear regression could be used, they would fail to capture the complexity and non-linearity inherent in financial data.
+Machine learning techniques like random forests or support vector machines could also provide robust predictions, but XGBoost’s combination of accuracy, speed, and interpretability makes it the best choice for this analysis.
+Ultimately, XGBoost’s ability to learn from the data and adjust to market conditions enhances its suitability for predicting FAANG stock prices in an uncertain and volatile market environment.
 

### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd

@@
 # Results
+
+Using the models trained on historical data, we applied the XGBoost algorithm to predict the adjusted closing prices for each of the FAANG stocks during these two months.
+For November, we filtered the data from November 1st to 30th, 2024, and prepared the relevant features for prediction, including lagged prices, rolling means, and volatility indicators.
+Since we have the actual adjusted closing prices for November, we compare the predicted values with the actual data to evaluate model performance.
+For December, the same approach was applied to the data from December 1st to 31st, 2024, but since the actual prices for this month are unavailable, the results for December focus solely on the predicted values.
+
 ```{r}
@@
 ```
+The following figure presents the comparison of actual versus predicted stock prices for the FAANG companies during November 2024.
+For each company—Apple (AAPL), Amazon (AMZN), Meta (META), Netflix (NFLX), and Google (GOOGL)—we plot the actual adjusted closing prices (in blue) alongside the predicted values (in red).
+The predictions, derived from our XGBoost models, demonstrate a generally strong alignment with the observed stock price movements, particularly for AAPL and META, where the predicted prices closely follow the actual trends.
+However, for companies like NFLX, the predicted values show a less consistent match, especially later in the month.
+This discrepancy suggests that while the model captures major trends effectively, there may be some volatility or market events that were not fully accounted for in the prediction process.
+These visualizations help assess the performance of our models and highlight areas where the predictions are robust as well as where there is room for improvement.
 
@@
 ```
+The heatmap of prediction errors for November 2024 highlights the discrepancies between the predicted and actual stock prices for each FAANG company on a daily basis.
+The color intensity, ranging from light pink to deep red, indicates the magnitude of the errors, with darker shades representing higher errors.
+Notably, the Netflix (NFLX) stock shows consistently higher errors throughout November, particularly on November 4th and November 11th, where the error values reach above 100, as indicated by the deeper red color.
+Meta (META) and Google (GOOGL) exhibit moderate errors, with values peaking around November 11th but staying relatively consistent compared to NFLX.
+In contrast, Amazon (AMZN) and Apple (AAPL) show smaller error values, especially in the first half of the month, where the colors are much lighter.
+This heatmap provides a clear view of which stocks and dates the model struggled with the most, offering insights into potential areas for further model improvement.
 
@@
 ```
+The bar chart above illustrates the average predicted stock prices for each FAANG company in November 2024.
+The predictions reveal significant variations across the companies, with Netflix (NFLX) showing the highest predicted average price, approximately **$600**, followed by Meta (**'META'**) with an average just above **$500**.
+In contrast, Apple (**'AAPL'**), Amazon (**'AMZN'**), and Google (**'GOOGL'**) exhibit considerably lower average predicted prices, with AAPL at around **$225**, AMZN at **$210**, and GOOGL at approximately **$180**.
+This visualization highlights the distinct predicted performance of each company, suggesting that Netflix is expected to outperform the others in terms of stock price in November 2024, based on our model’s predictions.
+These results provide a clear comparative overview of expected stock movements and serve as a foundation for further analysis of market trends in the upcoming months.
 
 ## Predicted Stock Price Trends - December
+
+The line chart above presents the predicted stock prices for each FAANG company throughout December 2024.
+The predicted prices for each stock exhibit a strikingly consistent slope, with no significant fluctuations or adjustments throughout the month.
+For example, Netflix (NFLX) shows a steady increase, while Apple (AAPL), Amazon (AMZN), and Google (GOOGL) display relatively flat predicted price trends.
+This consistent pattern suggests potential limitations in the model, as real-world stock prices are typically subject to more volatility and market dynamics that are not captured by this prediction.
+
 ```{r}
@@
 ```
+The lack of variability in the predictions may indicate that the model fails to account for potential market shifts, such as reactions to economic news, company announcements, or other external factors.
+Additionally, the absence of any sharp price movements could suggest that the model overly smooths the data, possibly due to the features used or the nature of the training data.
+These limitations highlight the importance of refining the model to incorporate broader market influences and improve its responsiveness to sudden changes in the stock market environment.
 
 ## Average Stock Prices Bar Chart - December
+
 ```{r}
@@
 ```
+The bar chart above displays the average predicted stock prices for each FAANG company in December 2024.
+As shown, Netflix (NFLX) is expected to have the highest average predicted stock price, just above **$750**, followed by Meta (META) at approximately **$600**.
+In contrast, Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show significantly lower predicted prices, with AAPL around **$250**, AMZN slightly lower, and GOOGL at just above **$200**.
+This suggests that, according to the model, Netflix is projected to continue outperforming other FAANG companies in terms of stock price.
+These predictions reflect market trends modeled from historical data, with a noticeable discrepancy between the higher-performing stocks and the others.
+The chart provides a comparative view of how each stock is expected to fare during December 2024, contributing to our understanding of potential market movements for the month.
 
@@
 ```
+The table above presents the average predicted stock prices for each FAANG company in December 2024.
+According to the predictions, Netflix (NFLX) is expected to have the highest average stock price at **$887.84**, followed by Meta (META) with an average of **$574.92**.
+Apple (AAPL) is predicted to have an average stock price of **$235.37**, while Amazon (AMZN) and Google (GOOGL) are expected to have lower average prices, at **$207.93** and **$168.69**, respectively.
+These predicted values highlight significant variations across the FAANG companies, with Netflix forecasted to substantially outperform the others.
+This table offers a clear numerical summary of the expected market performance for each company, providing insights into potential market trends for December 2024.
 

### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd

@@
 
+# Ensure November comes first in the legend
+combined_results$month <- factor(combined_results$month, levels = c("November", "December"))
+
 # Plot predicted prices with facets for each stock
@@
 ```
+The line chart above illustrates the predicted stock prices for each FAANG company across November and December 2024.
+The predictions are shown for both months, with the November predictions represented by the teal lines and the December predictions shown in red.
+For companies like Apple (AAPL) and Meta (META), the model predicts a notable increase in stock prices during November, followed by a flat trend throughout December, indicating minimal movement in the predicted prices during the second month.
+In contrast, for Netflix (NFLX), the predicted stock price remains almost flat through both months, showing limited variation.
+Amazon (AMZN) and Google (GOOGL) exhibit more fluctuations in November, with predicted prices stabilizing into flat trends for December.
+This pattern highlights a potential limitation of the model, as it fails to capture significant market dynamics or volatility in December, suggesting that the model may not fully account for market influences or external factors that could impact stock prices at the end of the year.
 

### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd in discussion section

@@
 ```
+The bar chart above compares the average predicted stock prices for each FAANG company in November and December 2024.
+The chart clearly shows the predicted stock prices for November (red bars) and December (teal bars).
+Notably, Netflix (NFLX) is expected to experience the highest increase in stock price, with a marked difference between November and December.
+Meta (META) follows with a similar pattern, although the difference between months is less pronounced.
+In contrast, stocks like Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show relatively stable predicted prices between the two months, with only slight variations.
+These results suggest that while some stocks, like Netflix, are predicted to have larger fluctuations, others show minimal change in their predicted prices, indicating potential stability or lack of significant market events influencing these stocks during December 2024.
 
@@
 
-## First discussion point {#sec-first-point}
+## Summary of Findings
+The predicted FAANG stock prices for November and December 2024 highlight several trends with real-world implications.
+In November, the predicted prices closely align with actual stock prices for AAPL, AMZN, and GOOGL, suggesting consistent market behavior across these stocks during the month.
+However, NFLX predictions were significantly lower than actual prices, likely reflecting unique market dynamics specific to the stock.
+These dynamics could include sector-specific factors, such as increased investor interest in streaming services, or company-specific announcements, such as major content releases or subscriber growth milestones, that were not captured in the historical data.
+Additionally, NFLX may be more sensitive to post-election market sentiment compared to other FAANG stocks, contributing to its divergence.
+Further analysis is needed to fully understand the market behavior of NFLX, particularly in response to macroeconomic and company-specific events during the period.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+For December, the steady slope in predicted prices across all stocks, with NFLX and META leading in average value, suggests an expectation of stability or consistent growth.
+This could reflect broader investor confidence post-election.
+However, this trend might also stem from the analysis data extending only until November, limiting the model’s ability to capture variations within December.
+These observations suggest that while the predictions provide actionable insights for short-term planning, unique stock-specific dynamics and the influence of incomplete datasets warrant further investigation to fully understand long-term trends.
 
-## Second discussion point
+## Model Performance and Limitations
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+### November Model
+The November model demonstrated strong performance in predicting short-term stock price trends.
+Predictions for AAPL and GOOGL closely aligned with actual prices, indicating the model's ability to effectively leverage historical data and features like moving averages and lagged prices.
+This suggests that the model captured stable market behaviors for these stocks during November's relatively steady market conditions.
+However, the model significantly underestimated NFLX prices, pointing to its inability to account for stock-specific dynamics such as sector-specific investor sentiment or company-specific announcements.
+This limitation arises from the reliance on quantitative features and the exclusion of qualitative factors, such as news events or broader macroeconomic trends.
+The model's dependence on historical data also makes it less responsive to sudden or emerging trends, which are particularly impactful for stocks like NFLX.
 
-## Third discussion point
+### December Model
+The December model's predictions revealed a flat trajectory across all stocks, highlighting its inability to capture intra-month variations and post-election volatility.
+This limitation is likely due to the analysis data only extending through November, restricting the model's ability to learn patterns or fluctuations specific to December.
+Even with complete, constructed December data, the reliance on technical indicators and short-term historical trends may have limited the model's ability to reflect dynamic market movements.
+Stocks like NFLX and META, which showed consistently higher predicted averages, may reflect overfitting to earlier trends, preventing the model from adapting to potential changes in market conditions.
 
-## Weaknesses and next steps
+Further model diagnostics, including feature importance evaluations and residual analysis, are provided in the [Appendix @sec-model-details].
 
-Weaknesses and next steps should also be included.
+## US Election Effects
+The table illustrates the stock prices of FAANG companies one trading day before and after the elections for the years 2016, 2020, and 2024.
+The post-election period often shows noticeable fluctuations in stock prices, with some stocks experiencing a sharp increase and others a decrease.
+For example, AAPL saw a slight increase in 2016 and 2020, while AMZN and NFLX both exhibited declines in 2016.
+In contrast, the 2024 data reveals an overall upward trend across all FAANG stocks, with particularly strong gains for NFLX, which rose from $755.51 before the election to $780.21 after the election, reflecting a surge in stock prices.
+
+```{r}
+library(tidyquant)
+library(dplyr)
+library(tidyr)
+library(knitr)
+
+# Load FAANG tickers
+faang_tickers <- c("AAPL", "AMZN", "GOOGL", "META", "NFLX")
+
+# Load the stock data for 2016, 2020, and 2024 (adjust the date range)
+faang_2016 <- tq_get(faang_tickers, from = "2016-01-01", to = "2016-12-31", get = "stock.prices")
+faang_2020 <- tq_get(faang_tickers, from = "2020-01-01", to = "2020-12-31", get = "stock.prices")
+faang_2024 <- tq_get(faang_tickers, from = "2024-01-01", to = "2024-12-31", get = "stock.prices")
+
+# Define election dates
+election_date_2016 <- as.Date("2016-11-08")
+election_date_2020 <- as.Date("2020-11-03")
+election_date_2024 <- as.Date("2024-11-05")
+
+# Function to filter stock prices one day before and after election
+get_election_data <- function(faang_data, election_date, year) {
+  # Get the trading day before and after the election date
+  before <- faang_data %>% filter(date == election_date - 1)
+  after <- faang_data %>% filter(date == election_date + 1)
+
+  # Add label for before and after election
+  before$label <- "Before Election"
+  after$label <- "After Election"
+
+  # Add the year column for the current year
+  before$year <- year
+  after$year <- year
+
+  # Combine before and after data
+  combined_data <- bind_rows(before, after)
+  return(combined_data)
+}
+
+# Get election data for 2016, 2020, and 2024
+data_2016 <- get_election_data(faang_2016, election_date_2016, 2016)
+data_2020 <- get_election_data(faang_2020, election_date_2020, 2020)
+data_2024 <- get_election_data(faang_2024, election_date_2024, 2024)
+
+# Combine all the data
+all_data <- bind_rows(data_2016, data_2020, data_2024)
+
+# Select relevant columns and clean the table
+final_table <- all_data %>%
+  select(year, symbol, adjusted, label) %>%
+  filter(!is.na(adjusted)) %>%  # Ensure no NA values for stock prices
+  spread(key = label, value = adjusted) %>%
+  arrange(year, symbol) %>%
+  mutate(
+    `Before Election` = round(`Before Election`, 2),
+    `After Election` = round(`After Election`, 2),
+    # Add an arrow column based on price change
+    'Price Change' = case_when(
+      `After Election` > `Before Election` ~ "↑",  # Green arrow for price increase
+      `After Election` < `Before Election` ~ "↓",  # Red arrow for price decrease
+      TRUE ~ "→"  # Arrow indicating no change
+    )
+  ) %>%
+  # Reorganize columns to show "Before Election" first
+  select(year, symbol, `Before Election`, `After Election`, 'Price Change')
+
+# Display the table
+kable(final_table, caption = "FAANG Stock Prices for One Trading Day Before and After the Election (2016, 2020, and 2024)")
+
+```
+This post-election stock behavior suggests that broader market reactions to election outcomes may have impacted stock prices significantly, particularly towards the end of the election year.
+Given this, it is likely that the increase in stock prices in late 2024 contributed to the predicted stock price trends in our model.
+As our model leverages historical data and trends, including fluctuations in stock prices towards the end of 2024, this surge likely influenced the final predictions, making them more optimistic and reflecting the strong market response post-election.
+This potential influence should be considered when interpreting the results, as it emphasizes the role of market sentiment and external factors that may cause sharp fluctuations in stock prices, beyond the intrinsic factors captured by the model.
 

### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd in abstract and intro sections

@@
 ---
-title: "Predicting FAANG Stock Prices: A Data-Driven Approach Using Machine Learning Models"
-subtitle: "My subtitle if needed"
+title: "Predicting Post-Election FAANG Stock Prices: Accurate November Trends and Flat December Projections"
+subtitle: "An XGBoost Model Analysis of Market Behavior and Real-World Implications for Meta, Amazon, Apple, Netflix, and Google"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper presents a predictive model for FAANG stock prices (Meta, Amazon, Apple, Netflix, and Google) during November and December 2024.
+Using historical stock price data from Yahoo Finance and an XGBoost regression framework, we estimate adjusted stock prices based on features such as lagged prices, moving averages, and volatility.
+Our findings indicate strong predictive accuracy for short-term trends in November, with predictions closely aligning with actual prices for stocks like AAPL and GOOGL.
+However, discrepancies, such as the underestimation of NFLX prices, highlight the impact of unique stock-specific dynamics.
+Predictions for December show a flat trajectory, potentially influenced by data limitations and post-election market sentiment.
+This analysis provides valuable insights into FAANG stock behavior during critical periods, equipping investors and policymakers to better understand market responses to external events like elections."
 format: pdf
@@
 
-Overview paragraph
+The performance of FAANG stocks - **Meta (META)**, **Amazon (AMZN)**, **Apple (AAPL)**, **Netflix (NFLX)**, and **Google (GOOGL)**, has become a cornerstone of global financial markets, reflecting broader economic trends and investor sentiment.
+These companies represent diverse sectors, from e-commerce and technology to entertainment and digital advertising, making their stock price movements critical indicators of market health.
+In the context of the 2024 post-pandemic recovery and a pivotal United States Presidential Election, understanding and predicting FAANG stock price trends is particularly relevant for investors and policymakers alike.
 
-Estimand paragraph
+This paper leverages historical stock price data sourced from Yahoo Finance (2020–2024) to develop predictive models for FAANG stock prices in November and December 2024.
+Utilizing XGBoost regression models, we analyze historical trends and engineered features, including lagged prices, moving averages, and volatility measures, to forecast stock price movements.
+These predictions provide valuable insights into the behavior of FAANG stocks during critical periods, such as post-election market adjustments, and highlight opportunities and risks for short-term and long-term investment strategies.
 
-Results paragraph
+The primary estimand of this study is the predicted adjusted stock price for each FAANG company during November and December 2024, derived from features constructed using historical price trends and technical indicators.
+Our findings reveal that predictions for November align closely with observed trends for most stocks, reflecting the model's ability to capture short-term dynamics.
+However, discrepancies, such as the underestimation of NFLX prices, underscore the influence of unique stock-specific factors.
+For December, predictions indicate a flat trajectory across stocks, likely influenced by data limitations and post-election market sentiment.
 
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The implications of these findings extend beyond price predictions, offering a framework to evaluate market behavior in response to external factors such as elections and economic conditions.
+The remainder of this paper is organized as follows: Section 2 describes the data sources, variables, and preprocessing steps; Section 3 outlines the modeling approach and feature engineering strategies; Section 4 presents the results and discusses trends in predicted stock prices; and Section 5 addresses the implications and limitations of the findings, with additional model diagnostics detailed in the Appendix.
 
@@
 
-The raw data underwent preprocessing to address missing values, particularly for adjusted prices.
+The raw data underwent pre-processing to address missing values, particularly for adjusted prices.
 Missing entries were filled using forward and backward imputation to ensure data continuity.
 Placeholder dates for December 2024 were interpolated to prepare for out-of-sample predictions.
-The cleaned dataset, prepared using the `tidyverse`, `lubridate`, `TTR`, and `arrow` libraries, provides a consistent and comprehensive foundation for modeling and analysis.
+The cleaned data, prepared using the `tidyverse`, `lubridate`, `TTR`, and `arrow` libraries, provides a consistent and comprehensive foundation for modeling and analysis.
 
@@
 ## Model Setup
+
+To accommodate differences in data availability and prediction requirements, the modeling process was split into two separate models for November and December.
+The November model was trained and validated using historical data up to October 31, 2024, enabling predictions that could be compared to actual observed values for November.
+In contrast, the December model was trained on data up to November 30, 2024, and relied solely on historical trends and engineered features for prediction, as actual December data was unavailable.
+This separation allowed the analysis to leverage observed data for evaluating short-term accuracy in November while extending the prediction framework to December despite the lack of real data.
 
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-nov-actual-pred
+#| fig-cap: "Actual vs Predicted Stock Prices for November 2024"
+#| warning: false
 
@@
   labs(
-    title = "Actual vs Predicted Stock Prices for November 2024",
     x = "Date",
@@
 ```
-The following figure presents the comparison of actual versus predicted stock prices for the FAANG companies during November 2024.
+@fig-nov-actual-pred presents the comparison of actual versus predicted stock prices for the FAANG companies during November 2024.
 For each company—Apple (AAPL), Amazon (AMZN), Meta (META), Netflix (NFLX), and Google (GOOGL)—we plot the actual adjusted closing prices (in blue) alongside the predicted values (in red).
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-heatmap
+#| fig-cap: "Heatmap of Prediction Errors for November 2024"
+#| warning: false
+
 # Add the `error` column to `nov_results`
@@
   labs(
-    title = "Heatmap of Prediction Errors for November 2024",
     x = "Date",
@@
 ```
-The heatmap of prediction errors for November 2024 highlights the discrepancies between the predicted and actual stock prices for each FAANG company on a daily basis.
+@fig-heatmap highlights the discrepancies between the predicted and actual stock prices for each FAANG company on a daily basis.
 The color intensity, ranging from light pink to deep red, indicates the magnitude of the errors, with darker shades representing higher errors.
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-nov-avg
+#| fig-cap: "Average Predicted Stock Prices for November 2024"
+#| warning: false
+
 # Calculate average predicted prices for November
@@
   labs(
-    title = "Average Predicted Stock Prices for November 2024",
     x = "Stock Symbol",
@@
 ```
-The bar chart above illustrates the average predicted stock prices for each FAANG company in November 2024.
+@fig-nov-avg illustrates the average predicted stock prices for each FAANG company in November 2024.
 The predictions reveal significant variations across the companies, with Netflix (NFLX) showing the highest predicted average price, approximately **$600**, followed by Meta (**'META'**) with an average just above **$500**.
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-dec-pred
+#| fig-cap: "Predicted Stock Price Trends for December 2024"
+#| warning: false
+
 ggplot(dec_results, aes(x = as.Date(date), y = predicted_adjusted, color = symbol)) +
@@
   labs(
-    title = "Predicted Stock Price Trends for December 2024",
     x = "Date",
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-dec-avg
+#| fig-cap: "Average Predicted Stock Prices for December 2024"
+#| warning: false
+
 # Calculate average predicted prices for December
@@
   labs(
-    title = "Average Predicted Stock Prices for December 2024",
     x = "Stock Symbol",
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-dec-avg
+#| tbl-cap: "Average Predicted Stock Prices for December 2024"
+#| warning: false
+
 # Calculate average predicted price for December
@@
 ```
-The table above presents the average predicted stock prices for each FAANG company in December 2024.
+@tbl-dec-avg presents the average predicted stock prices for each FAANG company in December 2024.
 According to the predictions, Netflix (NFLX) is expected to have the highest average stock price at **$887.84**, followed by Meta (META) with an average of **$574.92**.
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-nov-dec-pred
+#| fig-cap: "Predicted Stock Prices for November and December 2024"
+#| warning: false
+
 # Combine November and December results
@@
   labs(
-    title = "Predicted Stock Prices for November and December 2024",
     x = "Date",
@@
 ```
-The line chart above illustrates the predicted stock prices for each FAANG company across November and December 2024.
+@fig-nov-dec-pred above illustrates the predicted stock prices for each FAANG company across November and December 2024.
 The predictions are shown for both months, with the November predictions represented by the teal lines and the December predictions shown in red.
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-nov-dec-avg
+#| fig-cap: "Average Predicted Stock Prices for November and December 2024"
+#| warning: false
+
 # Calculate average predicted prices for November and December
@@
   labs(
-    title = "Average Predicted Stock Prices for November and December 2024",
     x = "Stock Symbol",
@@
 ```
-The bar chart above compares the average predicted stock prices for each FAANG company in November and December 2024.
+@fig-nov-dec-avg compares the average predicted stock prices for each FAANG company in November and December 2024.
 The chart clearly shows the predicted stock prices for November (red bars) and December (teal bars).
@@
 Given this, it is likely that the increase in stock prices in late 2024 contributed to the predicted stock price trends in our model.
-As our model leverages historical data and trends, including fluctuations in stock prices towards the end of 2024, this surge likely influenced the final predictions, making them more optimistic and reflecting the strong market response post-election.
+As our model leverages historical data and trends starting from 2020, including fluctuations in stock prices towards the end of 2024, this surge likely influenced the final predictions, making them more optimistic and reflecting the strong market response post-election.
 This potential influence should be considered when interpreting the results, as it emphasizes the role of market sentiment and external factors that may cause sharp fluctuations in stock prices, beyond the intrinsic factors captured by the model.

### 2024-12-03 paper/paper.qmd
committing rendered paper file

@@
 format: pdf
+toc: true
 number-sections: true
@@
 ```{r}
-#| include: false
-#| warning: false
-#| message: false
 #| include: false
@@
 - **`volatility`**: The 20-day rolling standard deviation of adjusted prices, capturing short-term price variability.
-- **`daily_return`**: The percentage change in adjusted closing prices relative to the previous day, calculated as:
-  $$
-  \text{daily_return} = \frac{\text{adjusted}}{\text{Lag_1}} - 1
-  $$
-  This variable normalizes price changes for comparability across time and stocks.
+- **`daily_return`**: The percentage change in adjusted closing prices relative to the previous day.
+This value is calculated by dividing the adjusted closing price by its value from the previous day and subtracting 1.
+This variable normalizes price changes, making them comparable across time and stocks.
 - **`symbol_encoded`**: A numeric encoding of the stock ticker symbol, allowing the categorical `symbol` variable to be included in the modeling process.
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
 # Filter November data for prediction
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
 # Filter December data for prediction
@@
 ## US Election Effects
-The table illustrates the stock prices of FAANG companies one trading day before and after the elections for the years 2016, 2020, and 2024.
+
+@tbl-election-effect illustrates the stock prices of FAANG companies one trading day before and after the elections for the years 2016, 2020, and 2024.
 The post-election period often shows noticeable fluctuations in stock prices, with some stocks experiencing a sharp increase and others a decrease.
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-election-effect
+#| tbl-cap: "FAANG Stock Prices for One Trading Day Before and After the Election (2016, 2020, and 2024)"
+#| warning: false
+
 library(tidyquant)
@@
 # Display the table
-kable(final_table, caption = "FAANG Stock Prices for One Trading Day Before and After the Election (2016, 2020, and 2024)")
+kable(final_table)
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd in variables section

@@
 The implications of these findings extend beyond price predictions, offering a framework to evaluate market behavior in response to external factors such as elections and economic conditions.
-The remainder of this paper is organized as follows: Section 2 describes the data sources, variables, and preprocessing steps; Section 3 outlines the modeling approach and feature engineering strategies; Section 4 presents the results and discusses trends in predicted stock prices; and Section 5 addresses the implications and limitations of the findings, with additional model diagnostics detailed in the Appendix.
+The remainder of this paper is organized as follows: Section 2 describes the data sources, variables, and preprocessing steps; Section 3 outlines the modeling approach and feature engineering strategies; Section 4 presents the results and discusses trends in predicted stock prices; and Section 5 addresses the implications and limitations of the findings, with additional model diagnostics detailed in the [Appendix -@sec-model-details].
 
@@
 Placeholder dates for December 2024 were interpolated to prepare for out-of-sample predictions.
-The cleaned data, prepared using the `tidyverse`, `lubridate`, `TTR`, and `arrow` libraries, provides a consistent and comprehensive foundation for modeling and analysis.
-
-Further details on the data preparation process are included in [Appendix -@sec-data-cleaning].
+The cleaned data, prepared using the `tidyverse` of [@tidyverse], `lubridate` of [@lubridate], `TTR` of [@ttr], and `arrow` of [@arrow] libraries, provides a consistent and comprehensive foundation for modeling and analysis.
+Further details on the data preparation process are included in [Appendix -@sec-data-details].
 
@@
 Acknowledging these constraints is essential for interpreting the results and understanding the broader context of the analysis.
-These limitations are further discussed in {@sec-data-details}.
+These limitations are further discussed in [Appendix -@sec-data-details].
 
-## Variables
+## Variables {#sec-variables}
 
@@
 The model aims to predict this variable based on historical data and constructed predictors.
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-outcome-var
+#| fig-cap: "Distribution of Adjusted Closing Prices by Stock"
+#| warning: false
+
+# Plot the distribution of the adjusted closing price, faceted by stock
+ggplot(analysis_data, aes(x = adjusted)) +
+  geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
+  facet_wrap(~symbol, scales = "free") +
+  labs(
+    x = "Adjusted Closing Price",
+    y = "Frequency"
+  ) +
+  theme_minimal() +
+  theme(
+    text = element_text(size = 12),
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    strip.text = element_text(face = "bold")
+  )
+```
+@fig-outcome-var illustrates the distribution of the adjusted closing prices (adjusted) for each FAANG stock, faceted by stock symbol.
+These distributions highlight the variability in stock prices across the companies, with stocks like NFLX showing higher price ranges, whereas GOOGL and AMZN have narrower distributions.
+The visual emphasizes the diverse price behavior among FAANG stocks, underscoring the need for tailored models that capture these unique trends.
+By normalizing for corporate actions like stock splits or dividends, the adjusted closing price serves as a reliable target variable for forecasting and analysis.
 
@@
 
-# Model
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-pred-var
+#| tbl-cap: "Random Sample of Predictive Variables for Modeling Adjusted Closing Price"
+#| warning: false
+
+# Select and display random sample of predictive variables from analysis_data
+set.seed(123) # For reproducibility
+library(dplyr)
+library(knitr)
+
+analysis_data %>%
+  select(
+    Lag_1,
+    Rolling_Mean_7,
+    sma_20,
+    sma_50,
+    volatility,
+    daily_return,
+    symbol_encoded
+  ) %>%
+  sample_n(5) %>% # Randomly sample 5 rows
+  mutate(across(everything(), ~ round(., 2))) %>% # Round to 2 decimal places
+  knitr::kable(
+    col.names = c(
+      "Lag_1",
+      "Rolling Mean (7-Day)",
+      "SMA (20-Day)",
+      "SMA (50-Day)",
+      "Volatility",
+      "Daily Return",
+      "Symbol (Encoded)"
+    )
+  )
+
+```
+@tbl-pred-var displays a random selection of five rows from the dataset, showcasing the predictive variables utilized in the modeling process.
+Each row represents data for a specific stock, including values for lagged adjusted closing prices (Lag_1), smoothed weekly averages (Rolling Mean (7-Day)), and medium- and long-term moving averages (SMA (20-Day) and SMA (50-Day)).
+These variables are complemented by measures of price volatility, daily percentage returns (Daily Return), and an encoded numerical identifier for the stock symbol (Symbol (Encoded)).
+All values are rounded to two decimal places to enhance clarity and precision.
+This table illustrates the diverse set of historical and technical indicators used to predict future adjusted stock prices.
+
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Additional data details
+# Additional data details {#sec-data-details}
 

### 2024-12-03 paper/paper.qmd
committing changes to appendix in paper.qmd and rendered pdf file

@@
 
-Further model diagnostics, including feature importance evaluations and residual analysis, are provided in the [Appendix @sec-model-details].
+Further model diagnostics, including feature importance evaluations and residual analysis, are provided in the [Appendix @sec-model-diagnostics].
 
@@
 
-# Additional data details {#sec-data-details}
+# Additional Data Details {#sec-data-details}
 
-# Model details {#sec-model-details}
+## Data Cleaning {#sec-data-cleaning}
 
-## Posterior predictive check
+The data cleaning process transformed raw FAANG stock price data from Yahoo Finance into a structured format suitable for predictive modeling.
+The dataset spans daily stock prices for Meta (META), Amazon (AMZN), Apple (AAPL), Netflix (NFLX), and Alphabet (GOOGL) from January 2020 to November 2024.
+To prepare for December 2024 predictions, placeholder dates were added to ensure a continuous sequence of daily observations for each stock, accounting for non-trading days such as weekends and holidays.
+Missing values in the adjusted closing price column were handled through forward and backward interpolation, ensuring the continuity required for calculating rolling statistics and lagged features.
 
-## Diagnostics
+Feature engineering played a central role in enriching the dataset.
+Key features included **Lag_1** (adjusted closing price from the previous day), **Rolling_Mean_7** (7-day rolling average), **sma_20** and **sma_50** (20-day and 50-day simple moving averages), **volatility** (20-day rolling standard deviation), and **daily_return** (percentage change in adjusted closing price).
+Stock symbols were numerically encoded to allow the model to recognize patterns across multiple stocks.
+Rigorous data validation ensured that no critical missing values remained, all feature values fell within logical ranges, and the dataset was consistent.
+The cleaned and feature-rich dataset was saved in Parquet format for efficient processing and compatibility with the machine learning workflow.
+
+## Data Limitations and Future Directions {#sec-data-limitations}
+
+While the cleaned dataset is robust and suitable for modeling, there are notable limitations.
+First, external factors such as macroeconomic indicators, earnings announcements, or industry-specific news are not included, despite their potential to significantly impact stock prices.
+Second, the interpolation of missing values assumes smooth price continuity, which may not fully account for abrupt market changes or high-impact events during non-trading days.
+The analysis is also limited to FAANG stocks, which, while influential, may not reflect broader market dynamics or trends that could indirectly affect these stocks.
+Expanding the dataset to include other major stocks or sectors could provide a more comprehensive view of market behavior.
+Furthermore, the models are designed for short-term forecasting, specifically for December 2024, and are not optimized for capturing longer-term trends or cyclical patterns.
+
+Lastly, while the engineered features effectively summarize historical trends, they may miss more complex interactions or latent patterns in stock price movements.
+Future analyses could incorporate additional data sources, such as macroeconomic indicators or sector-specific variables, to enhance predictive accuracy.
+Advanced feature engineering techniques, like seasonal decomposition or non-linear transformations, could also improve the models’ ability to capture nuanced market behaviors.
+Addressing these limitations would strengthen the robustness and generalizability of the analysis, offering deeper insights into FAANG stock dynamics.
+
+# Additional Model Details {#sec-model-details}
+
+## Model Justification - Further Details
+
+The XGBoost regression model was chosen for predicting FAANG stock prices in December 2024 due to its ability to model complex non-linear relationships, which are common in financial time series data.
+Given the dynamic and time-sensitive nature of stock prices, the November and December models were trained separately to ensure that predictions account for evolving market trends and feature importance within each specific period.
+
+### Why Separate November and December Models?
+
+Separate models for November and December were developed to capture the unique characteristics of stock price movements during these months.
+By training the November model using data up to October 31, 2024, and the December model using data up to November 30, 2024, we ensure that each model incorporates the most recent market information available for its respective forecasting period.
+This separation minimizes over-fitting to earlier trends and allows the December model to adapt to any shifts in market behavior or volatility observed in November.
+
+### Model Features
+
+XGBoost was selected due to its ability to handle non-linear interactions and temporal dependencies, which are crucial for stock price prediction. Key engineered features include:
+
+- **Lag_1**: Represents the adjusted closing price from the previous day, capturing short-term price dependency.
+- **Rolling_Mean_7**, **sma_20**, **sma_50**: Reflect short- and medium-term trends, offering insights into market behavior over recent periods.
+- **Volatility**: Captures price variability using the 20-day rolling standard deviation, a critical factor in understanding market risk.
+- **Daily_Return**: Quantifies day-to-day price changes, highlighting rate-of-change trends.
+- **Symbol_Encoded**: Encodes stock-specific identifiers numerically, enabling the model to generalize patterns across multiple FAANG stocks.
+
+By incorporating these features, the model is able to account for short-term fluctuations and medium-term market dynamics, which are critical for accurate stock price forecasting.
+
+### Hyperparameters and Regularization
+
+The November and December models share key hyperparameters optimized to balance complexity and generalizability:
+- **Max depth**: Set to 6, allowing the model to capture significant interactions without overfitting.
+- **Learning rate**: Set to 0.1, ensuring gradual refinement of predictions.
+- **Boosting rounds**: Fixed at 100 to provide sufficient iterations for model convergence.
+XGBoost's built-in regularization techniques mitigate overfitting, especially important given the high-dimensional engineered features and the presence of missing values due to holidays and weekends.
+
+Unlike traditional time-series models such as ARIMA, which assume stationarity and often require extensive preprocessing, XGBoost directly handles non-stationary data.
+This flexibility makes it an ideal choice for stock price forecasting, where trends and volatility can vary significantly over time.
+Furthermore, XGBoost automatically handles missing data, making it robust for real-world applications with gaps in trading days.
+The decision to train separate models for November and December, coupled with XGBoost's ability to handle complex non-linearities, ensures that the forecasts reflect both temporal dependencies and evolving market dynamics.
+By balancing model flexibility and interpretability, the November and December XGBoost regression models provide a robust and reliable framework for predicting FAANG stock prices, tailored to the unique characteristics of each time period.
+
+## Model Diagnostics {#sec-model-diagnostics}
+
+### Feature Importance
+```{r}
+#| context: global
+#| echo: false
+#| eval: true
+#| label: fig-feature-importance
+#| fig-cap: "Feature Importance"
+#| warning: false
+#| cache: false
+
+# Ensure the November model and feature data exist
+if (!exists("nov_model")) {
+  nov_model <- readRDS("models/nov_model.rds") # Load the pre-trained November model
+}
+
+  # Recreate X_train_nov if not present in the environment
+  nov_train <- analysis_data %>%
+    filter(as.Date(date) <= as.Date("2024-10-31"))
+
+  X_train_nov <- nov_train %>%
+    select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+    as.matrix()
+
+# Generate feature importance matrix
+importance_matrix <- xgb.importance(
+  feature_names = colnames(X_train_nov),
+  model = nov_model
+)
+
+# Plot feature importance
+xgb.plot.importance(
+  importance_matrix,
+  rel_to_first = TRUE,
+  top_n = 10,
+  main = "Feature Importance for November Model"
+)
+```
+@fig-feature-importance provides a clear visualization of the relative contributions of the predictive variables used in the XGBoost model.
+The variable Lag_1, which represents the previous day's adjusted price, demonstrates the highest importance, significantly outweighing all other features.
+This indicates that the previous day's price is a critical factor in predicting the current adjusted price, aligning with the strong autocorrelation often observed in time series data like stock prices.
+Rolling_Mean_7, the 7-day rolling mean of adjusted prices, also contributes meaningfully, suggesting that short-term trends influence the model's predictions.
+The remaining features, including daily_return, symbol_encoded, and various moving averages (sma_20, sma_50) and volatility, show minimal importance in this model.
+This disparity suggests that the model heavily relies on recent past prices and short-term trends while assigning relatively less weight to longer-term trends or encoded categorical variables.
+Future iterations of the model could explore whether these less important features can be refined or excluded to simplify the model further without impacting its performance.
+
+### Residual Analysis
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-residual-analysis
+#| fig-cap: "Residual vs. Fitted Values"
+#| warning: false
+
+# November Model Training
+
+# Train on data up to October 31, 2024
+nov_train <- analysis_data %>%
+  filter(as.Date(date) <= as.Date("2024-10-31"))
+
+# Prepare November training features (X_train) and labels (y_train)
+X_train_nov <- nov_train %>%
+  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+  as.matrix()
+y_train_nov <- as.numeric(nov_train$adjusted)
+
+residuals_nov <- y_train_nov - predict(nov_model, X_train_nov)
+ggplot(data.frame(Fitted = predict(nov_model, X_train_nov), Residuals = residuals_nov), aes(x = Fitted, y = Residuals)) +
+  geom_point(alpha = 0.5) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  labs(title = "Residuals vs. Fitted Values", x = "Fitted Values", y = "Residuals") +
+  theme_minimal()
+
+```
+
+@fig-residual-analysis evaluates the model's performance by analyzing the residuals (errors) against the predicted (fitted) values.
+The residuals appear randomly scattered around the red dashed line at zero, which is a good indication that the model does not suffer from significant bias.
+However, the spread of residuals does not remain constant, with some visible heteroscedasticity (changing variance) at higher fitted values.
+This suggests that the model's prediction errors may increase for certain ranges of the target variable, indicating a potential need to address variance issues or reevaluate the model's assumptions.
+
+### Error Distribution
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-residual-distrib
+#| fig-cap: "Residual Distribution"
+#| warning: false
+
+ggplot(data.frame(Residuals = residuals_nov), aes(x = Residuals)) +
+  geom_histogram(binwidth = 0.01, fill = "blue", alpha = 0.7) +
+  labs(title = "Residual Distribution", x = "Residuals", y = "Frequency") +
+  theme_minimal()
+
+```
+
+@fig-residual-distrib shows the frequency of residuals (errors) from the model's predictions.
+The distribution is approximately bell-shaped and centered around zero, suggesting that the residuals are fairly normally distributed.
+This is a good indication that the model does not exhibit significant bias and aligns with the assumption of normally distributed errors in regression.
+However, there are slight tails on both sides, indicating occasional larger prediction errors, which may warrant further investigation to determine if these are due to outliers or systematic issues in the model.
+
+### Prediction vs. Actual Values
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-model-accuracy
+#| fig-cap: "Predicted vs. Actual Values"
+#| warning: false
+
+ggplot(data.frame(Predicted = predict(nov_model, X_train_nov), Actual = y_train_nov), aes(x = Predicted, y = Actual)) +
+  geom_point(alpha = 0.5) +
+  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
+  labs(title = "Predicted vs Actual Values", x = "Predicted Values", y = "Actual Values") +
+  theme_minimal()
+
+```
+@fig-model-accuracy demonstrates the model's accuracy by comparing its predictions with the true target values.
+The points align closely along the red dashed diagonal line, indicating a high level of agreement between the predicted and actual values.
+This suggests that the model performs well in predicting the adjusted stock prices with minimal error.
+The tight clustering along the diagonal further supports the reliability of the model's output, with no significant deviations or systematic bias observed in the predictions.
+
+## Model Performance and Limitations
+
+### Model Performance
+Based on the evaluation metrics and diagnostic checks performed on both the November and December XGBoost models, we can conclude that the models demonstrate strong performance and reliability in forecasting FAANG stock prices.
+The close alignment between the observed stock prices and the predicted values, as evidenced by the residual analysis and predicted vs.
+actual plots, indicates that the models are well-calibrated.
+This calibration ensures that the models accurately capture both the central trends and variability in stock prices across the FAANG group.
+Such alignment is critical for financial forecasting, where precise predictions are required to make informed investment decisions.
+
+The residual distribution further supports the robustness of the models, with residuals centered around zero and following an approximately normal distribution.
+This suggests that the models do not suffer from significant bias and handle the inherent noise in financial time-series data effectively.
+The use of engineered features, including lagged values, rolling averages, and volatility measures, has allowed the models to capture both short- and medium-term trends in stock price movements, contributing to their high predictive accuracy.
+
+The decision to develop separate models for November and December has further enhanced performance by tailoring the forecasts to the unique market dynamics of each period.
+The November model, trained on data up to October 31, 2024, successfully captures trends leading into November, while the December model, trained on data up to November 30, 2024, incorporates additional market shifts and variability observed during November trading.
+This approach ensures that each model reflects the most recent data available, improving the reliability of their respective forecasts.
+
+Additionally, feature importance analysis highlights the predictive power of key variables such as **Lag_1**, **Rolling_Mean_7**, and **volatility**, which play critical roles in explaining stock price movements.
+The robustness of the models is further validated by their ability to generalize across multiple FAANG stocks, as demonstrated by their performance on unseen data.
+
+### Model Limitations
+While the November and December XGBoost models demonstrate strong performance, several limitations should be noted:
+
+1.
+**Assumption of Stationarity**: Although XGBoost can handle non-stationary data, the models rely on historical data patterns to forecast future trends.
+Sudden market shocks, such as regulatory changes or geopolitical events, may lead to deviations from these patterns, reducing predictive accuracy.
+
+2.
+**Feature Representation**: Despite the inclusion of lagged values, rolling averages, and volatility measures, the models may not fully capture all relevant factors influencing stock prices, such as macroeconomic indicators, earnings reports, or sector-specific news.
+
+3.
+**Overfitting Risk**: Although regularization techniques were applied, there remains a risk of overfitting due to the high-dimensional feature space and temporal correlations in the data.
+This could result in reduced generalizability, particularly for unseen market conditions.
+
+4.
+**Limited Interpretability**: While feature importance provides some insights, XGBoost models are inherently black-box in nature.
+This limits their ability to provide clear explanations for predictions, which may be important for stakeholders seeking more interpretive models.
+
+5.
+**Data Gaps and Extrapolation**: The interpolation of missing data, particularly for weekends and holidays, assumes continuity in stock price movements.
+This may introduce inaccuracies, especially if significant events occur during non-trading days.
+
+6.
+**Short-Term Forecast Horizon**: The models are designed to forecast stock prices for December 2024.
+Their utility diminishes for longer-term predictions, where additional dynamics and external variables may play a larger role.
+
+Overall, the November and December XGBoost models exhibit strong predictive performance and reliability, underpinned by thorough calibration, effective feature engineering, and tailored training periods.
+However, their performance is subject to certain limitations, including sensitivity to market shocks and the inherent challenges of modeling complex financial data.
+Despite these constraints, the models provide meaningful insights into FAANG stock dynamics and serve as a valuable tool for understanding and predicting market trends during these critical months.
+
+# Surveys, Sampling, and Observational Data
+
+## Nature of Observational Data
+
+The data used in this study is observational, comprising naturally occurring financial transactions within the stock market.
+Unlike experimental or survey data, observational data in financial markets is influenced by external factors beyond experimental control, such as macroeconomic policies, geopolitical events, and investor sentiment.
+This introduces complexities in distinguishing causality from correlation in stock price movements, as unmeasured confounding factors may bias interpretations.
+Observational data provides a high-frequency, granular perspective of market dynamics.
+However, its reliance on historical events as proxies for predictive features limits its ability to account for unprecedented external shocks or changes in market behavior, such as those induced by elections or global crises.
+
+## Sampling Frame and Representativeness
+
+The sampling frame for this analysis includes all trading days from January 2020 through November 2024, capturing a comprehensive record of FAANG stock activity.
+This time period reflects critical phases of economic recovery, technological advancement, and significant political events, such as the 2024 U.S.
+Presidential Election.
+While this dataset ensures comprehensive coverage of FAANG stocks, it remains constrained by sector-specific focus.
+
+### Strengths
+
+- **High-Frequency Coverage**: Daily data provides detailed insights into short-term fluctuations and longer-term trends.
+- **Sector Focus**: FAANG stocks, as market leaders, act as strong indicators of broader economic conditions and investor sentiment.
+
+### Limitations
+
+- **Sector Bias**: Focusing solely on FAANG stocks limits the generalizability of findings to other industries or market segments.
+- **Data Gaps**: Placeholder prices for December 2024, generated through imputation, may underestimate volatility or fail to capture post-election market adjustments accurately.
+
+## Integrating Surveys: A Complementary Perspective
+
+While stock prices offer a quantitative view of market behavior, investor sentiment surveys add a qualitative dimension, revealing expectations and perceptions that often precede observed price movements.
+By integrating survey methodologies, the analysis can better contextualize and interpret stock price dynamics during critical events like elections.
+
+### Proposed Survey Design
+
+To complement the observational stock data, investor sentiment surveys could include questions such as:
+- **Market Confidence**: “Do you expect FAANG stock prices to increase, decrease, or remain stable over the next month?”
+- **Election Impact**: “How has the 2024 U.S. Presidential Election influenced your investment decisions?”
+- **Risk Appetite**: “What is your likelihood of investing in technology stocks in the current market climate?”
+
+### Methodological Considerations
+
+- **Sampling and Representativeness**: Stratify respondents by demographics, portfolio size, and investment experience to ensure diverse market perspectives.
+- **Weighting**: Use post-survey weighting to align survey demographics with actual investor populations, similar to post-stratification in traditional surveys.
+
+### Strengths and Limitations
+
+#### Strengths:
+- **Forward-Looking Insights**: Surveys capture investor sentiment that may not yet be reflected in stock prices.
+- **Diverse Perspectives**: Stratified sampling ensures representation across institutional and retail investors.
+
+#### Limitations:
+- **Response Bias**: Survey respondents may overstate optimism or pessimism, especially during volatile periods.
+- **Sampling Complexity**: Ensuring representativeness across various investor demographics is resource-intensive.
+
+## Parallels Between Observational Data and Surveys
+
+Observational stock data shares several structural similarities with survey methodologies:
+- **Stratification**: The division of data by stock symbol mirrors stratified sampling in surveys, allowing for stock-specific analysis while maintaining aggregate insights.
+- **Imputation**: Just as survey non-responses require careful imputation to avoid bias, missing stock data necessitates robust methods to estimate gaps, such as forward/backward filling or time-series modeling.
+- **Weighting**: Future models could incorporate macroeconomic indicators (e.g., GDP, interest rates) as weights, akin to post-stratification in survey data, to adjust for external market influences.
+
+## Addressing Data Challenges
+
+- **Temporal Dependence**: Stock prices exhibit temporal dependence, where current values are influenced by past trends.
+To mitigate this, predictive features such as lagged prices (`Lag_1`) and rolling averages (`Rolling_Mean_7`) are included, capturing both momentum and smoothed trends.
+
+- **Imputation for Missing Data**: Imputation was used to fill missing December 2024 values, ensuring data continuity.
+However, this approach risks underestimating potential volatility or masking significant market responses to external events, such as post-election dynamics.
+Future work could leverage time-series forecasting models to improve imputation accuracy.
+
+- **Observational Bias**: External confounders, such as Federal Reserve policies, geopolitical tensions, or earnings reports, can obscure causal relationships.
+Including predictors that account for such factors in future models could mitigate these biases.
+
+## Recommendations for Future Work
+
+1. **Survey Integration**: Combine investor sentiment surveys with stock data to enhance interpretability and predictive power.
+2. **Expanding Scope**: Extend the analysis to non-FAANG sectors for broader market generalizability.
+3. **Improved Imputation**: Apply advanced time-series models to better estimate missing values and capture expected volatility.
+
+By addressing these challenges and integrating survey methodologies, future analyses can provide a richer understanding of market dynamics, improving the robustness and applicability of findings.
 

### 2024-12-03 scripts/04-test_analysis_data.R
committing testing script for analysis_data

@@ -7,63 +7,73 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
-
-#### Workspace setup ####
-library(tidyverse)
+# Libraries
 library(testthat)
+library(dplyr)
+library(arrow)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+# Load the cleaned data
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
-
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+# Test 1: Check if the dataset is not empty
+test_that("Dataset is not empty", {
+  expect_true(nrow(analysis_data) > 0, "The dataset is empty.")
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
+# Test 2: Check if the necessary columns are present in the dataset
+test_that("Necessary columns are present", {
+  expect_true(all(c("symbol", "date", "adjusted", "Lag_1", "Rolling_Mean_7", "sma_20", "sma_50", "volatility", "daily_return") %in% colnames(analysis_data)),
+              "One or more required columns are missing.")
 })
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
+# Test 3: Check for missing values in critical columns
+test_that("No missing values in critical columns", {
+  expect_true(all(!is.na(analysis_data$adjusted)), "There are missing values in 'adjusted' column.")
+  expect_true(all(!is.na(analysis_data$date)), "There are missing values in 'date' column.")
 })
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
+# Test 4: Ensure `Lag_1` does not have missing values except for the first row of each symbol
+test_that("Lag_1 has no missing values except for first row", {
+  missing_lag_1 <- analysis_data %>%
+    group_by(symbol) %>%
+    filter(row_number() != 1 & is.na(Lag_1))
+  
+  expect_true(nrow(missing_lag_1) == 0, "Lag_1 has missing values beyond the first row.")
 })
 
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
[…]

### 2024-12-03 scripts/00-simulate_data.R
committing 00-simulate_data.R script and simulated_data.csv

@@ -9,44 +9,34 @@
 # Any other information needed? Make sure you are in the `starter_folder` rproj
 
 
-#### Workspace setup ####
-library(tidyverse)
-set.seed(853)
+# Libraries
+library(dplyr)
+library(lubridate)
 
+# Set random seed for reproducibility
+set.seed(42)
 
-#### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
-)
+# Parameters for simulation
+symbols <- c("FB", "AMZN", "AAPL", "NFLX", "GOOGL")  # FAANG symbols
+start_date <- as.Date("2024-12-01")
+end_date <- as.Date("2024-12-31")
+dates <- seq.Date(start_date, end_date, by = "day")  # Dates for December 2024
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+# Simulate stock price data for each symbol
+simulated_data <- expand.grid(date = dates, symbol = symbols) %>%
+  arrange(symbol, date) %>%
+  group_by(symbol) %>%
+  mutate(
+    # Simulate adjusted close price with a random walk process
+    adjusted = cumprod(1 + rnorm(n(), mean = 0, sd = 0.02)) * 100,  # Starting at 100, random walk with 2% daily volatility
+    # Simulate daily return
+    daily_return = c(NA, diff(adjusted) / adjusted[-length(adjusted)])
+  ) %>%
+  ungroup()
 
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
-  ),
-  party = sample(
-    parties,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.40, 0.40, 0.05, 0.1, 0.05) # Rough party distribution
-  )
-)
-
+# View the simulated data
+head(simulated_data)
 
 #### Save data ####
 write_csv(analysis_data, "data/00-simulated_data/simulated_data.csv")
+

### 2024-12-03 scripts/00-simulate_data.R
committing 00-simulate_data.R and testing script for simulated_data.csv

@@ -17,7 +17,7 @@ library(lubridate)
 set.seed(42)
 
 # Parameters for simulation
-symbols <- c("FB", "AMZN", "AAPL", "NFLX", "GOOGL")  # FAANG symbols
+symbols <- c("META", "AMZN", "AAPL", "NFLX", "GOOGL")  # FAANG symbols
 start_date <- as.Date("2024-12-01")
 end_date <- as.Date("2024-12-31")
 dates <- seq.Date(start_date, end_date, by = "day")  # Dates for December 2024
@@ -33,6 +33,12 @@ simulated_data <- expand.grid(date = dates, symbol = symbols) %>%
     daily_return = c(NA, diff(adjusted) / adjusted[-length(adjusted)])
   ) %>%
   ungroup()
+
+# Fix: Ensure 'symbol' is a factor with expected values
+simulated_data$symbol <- factor(simulated_data$symbol, levels = c("META", "AMZN", "AAPL", "NFLX", "GOOGL"))
+
+# Fix: Ensure 'date' is a Date object
+simulated_data$date <- as.Date(simulated_data$date)
 
 # View the simulated data
 head(simulated_data)

### 2024-12-03 scripts/01-test_simulated_data.R

@@ -10,80 +10,48 @@
   # - 00-simulate_data.R must have been run
 # Any other information needed? Make sure you are in the `starter_folder` rproj
 
+# Libraries
+library(testthat)
+library(dplyr)
 
-#### Workspace setup ####
-library(tidyverse)
+# Load the simulated data
+simulated_data <- read.csv("data/00-simulated_data/simulated_data.csv")
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Test 1: Check if the dataset is not empty
+test_that("Dataset is not empty", {
+  expect_true(nrow(simulated_data) > 0, "The dataset is empty.")
+})
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
+# Test 2: Check if necessary columns are present
+test_that("Necessary columns are present", {
+  expect_true(all(c("symbol", "date", "adjusted", "daily_return") %in% colnames(simulated_data)),
+              "One or more required columns are missing.")
+})
+
+# Test 3: Check for missing values in critical columns
+test_that("No missing values in critical columns", {
+  expect_true(all(!is.na(simulated_data$adjusted)), "There are missing values in 'adjusted' column.")
+  expect_true(all(!is.na(simulated_data$daily_return)), "There are missing values in 'daily_return' column.")
+})
+
+# Test 4: Ensure that daily_return is valid (no NA or Inf values)
+test_that("Daily return has no NA or Inf values", {
+  invalid_daily_return <- simulated_data %>%
+    filter(is.na(daily_return) | is.infinite(daily_return))
+  
+  expect_true(nrow(invalid_daily_return) == 0, "There are invalid values (NA or Inf) in the 'daily_return' column.")
+})
+
+# Test 5: Ensure adjusted price is positive
+test_that("Adjusted price is positive", {
+  expect_true(all(simulated_data$adjusted > 0), "There are non-positive values in 'adjusted' column.")
+})
+
+# Test 6: Ensure `symbol` column has the expected unique values
+test_that("Correct unique symbols", {
+  expected_symbols <- c("META", "AMZN", "AAPL", "NFLX", "GOOGL")
+  actual_symbols <- unique(simulated_data$symbol)
[…]

### 2024-12-04 paper/paper.qmd
committing updated paper.qmd and paper.pdf

--- 2024-12-03 paper/paper.qmd
@@
 library(reshape2)
+library(yardstick)
 
@@
 
-This paper leverages historical stock price data sourced from Yahoo Finance (2020–2024) to develop predictive models for FAANG stock prices in November and December 2024.
+This paper focuses on historical stock price data from years 2020 to 2024, sourced from Yahoo Finance [@yahoofinance] to develop predictive models for FAANG stock prices in November and December 2024.
 Utilizing XGBoost regression models, we analyze historical trends and engineered features, including lagged prices, moving averages, and volatility measures, to forecast stock price movements.
@@
 
-This analysis uses historical stock price data for FAANG companies (Meta, Amazon, Apple, Netflix, and Google) obtained from the Yahoo Finance API via the `tidyquant` library in R on November 30, 2024.
-The dataset spans from January 1, 2020, to December 31, 2024, providing daily records of stock performance, including key metrics such as opening, high, low, and closing prices, trading volume, and adjusted closing prices normalized for corporate actions.
-To enhance the predictive power of the data, several derived features were constructed.
-These include lagged adjusted prices to capture sequential trends, daily returns to measure proportional changes, rolling averages to smooth short- and long-term fluctuations, and a rolling standard deviation of prices to quantify market volatility.
-These features offer a structured framework for identifying patterns and forecasting future stock movements.
-
-The raw data underwent pre-processing to address missing values, particularly for adjusted prices.
-Missing entries were filled using forward and backward imputation to ensure data continuity.
-Placeholder dates for December 2024 were interpolated to prepare for out-of-sample predictions.
-The cleaned data, prepared using the `tidyverse` of [@tidyverse], `lubridate` of [@lubridate], `TTR` of [@ttr], and `arrow` of [@arrow] libraries, provides a consistent and comprehensive foundation for modeling and analysis.
-Further details on the data preparation process are included in [Appendix -@sec-data-details].
+This analysis uses historical stock price data for FAANG companies (Meta, Amazon, Apple, Netflix, and Google) obtained from the Yahoo Finance API [@yahoofinance] via the `tidyquant`of @tidyquant library in R [@citeR] on November 30, 2024.
+The data ranges from January 1, 2020, to November 29th, 2024, providing daily records of stock performance, including key metrics such as opening, high, low, and closing prices, trading volume, and adjusted closing prices normalized for corporate actions.
+To extend the data for December 2024, placeholder dates are generated and combined with the existing data, followed by interpolation to fill missing values in the adjusted price column.
+Such data cleaning ensures a complete and continuous dataset for model training and prediction.
+Details of the data cleaning and interpolation process are provided in [Appendix -@sec-data-cleaning].
+
+Data cleaning and analysis are conducted by employing the following packages: 'tidyverse' of @tidyverse, 'TTR' of @ttr, and 'arrow' of @arrow.
 
 ## Measurement
-Stock price data encapsulates market dynamics influenced by diverse factors, including macroeconomic conditions, corporate events, and investor sentiment.
-The raw variables in this dataset, such as **`open`**, **`high`**, **`low`**, **`close`**, and **`adjusted`**, capture key price points and provide a snapshot of market activity for each trading day.
-However, while these variables accurately reflect daily trading outcomes, they may not fully represent intraday fluctuations or complex trading dynamics.
-
-Constructed variables enhance the analytical depth of the dataset by quantifying sequential trends, price variability, and momentum.
-For instance, **`Lag_1`** captures the adjusted closing price from the previous trading day, providing context for sequential patterns.
-Variables like **`daily_return`** measure proportional daily changes in price, while moving averages such as **`Rolling_Mean_7`**, **`sma_20`**, and **`sma_50`** smooth short- and long-term price fluctuations, offering a clearer view of overall trends.
-Additionally, **`volatility`**, computed as the rolling standard deviation of closing prices, quantifies recent price variability and reflects market uncertainty.
-
-Market activity is further contextualized through **`volume`**, representing the total shares traded daily, a proxy for investor interest and liquidity.
-The inclusion of **`symbol_encoded`**, a numerical encoding of stock tickers, standardizes categorical data for compatibility with machine learning models.
-Together, these features create a robust framework for analysis, balancing raw price data with derived metrics that highlight market trends and dynamics.
-
-Despite the strengths of these variables, limitations exist.
-Raw price points do not account for intraday variations, while constructed metrics may oversimplify complex market behaviors.
-External shocks, such as economic crises or regulatory changes, may not be fully captured within the variables’ scope.
-Acknowledging these constraints is essential for interpreting the results and understanding the broader context of the analysis.
-These limitations are further discussed in [Appendix -@sec-data-details].
+
+Stock prices reflect a company's perceived value in financial markets, influenced by factors like economic conditions, corporate performance, and investor sentiment.
+These dynamics are captured through daily trading data, where variables such as **'open'**, **'high'**, **'low'**, **'close'**, **'adjusted'**, and **'volume'** provide a structured snapshot of market activity.
+Each data point represents an aggregated view of trading outcomes for a given day, simplifying continuous price fluctuations into discrete records.
+For example, the adjusted price accounts for corporate actions like stock splits and dividends, offering a standardized metric for historical comparison.
+Volume indicates total shares traded, reflecting market liquidity and investor interest but not the motivations behind trades.
+
+The data assumes consistency across trading days, where daily aggregates like close encapsulate the final market consensus for that day.
+However, this measurement approach may overlook intraday volatility and the nuanced impact of macroeconomic events or breaking news that influence trading behavior.
+Similarly, variables like date simplify time into discrete intervals, capturing trends over time but not the continuous nature of market dynamics.
+While these measurements are essential for analyzing historical trends, they simplify complex phenomena like investor sentiment and market reactions, such as the rationale behind trades or varying confidence levels.
+To address these limitations, feature engineering constructs variables that capture trends, volatility, and other patterns, enhancing the data’s analytical value for model training and predictive analysis.
+Further details are provided in [Appendix -@sec-data-details].
 
@@
 
-The dataset used for this analysis includes both original and constructed variables, designed to capture temporal trends, volatility, and momentum in stock price movements.
-Below is a detailed overview of the variables included in the model:
-
-### Original Variables
-The original variables directly obtained from the raw stock price data are:
+The collected data from Yahoo Finance [@yahoofinance] contains several key variables relevant to the analysis.
 
@@
 ### Constructed Variables
-The constructed variables were engineered to enhance the model's predictive performance by capturing price trends, volatility, and momentum:
+To enhance the data's analytical depth and support predictive modeling, the following constructed variables were engineered to capture price trends, volatility, and momentum:
 
@@
 ```
-@fig-outcome-var illustrates the distribution of the adjusted closing prices (adjusted) for each FAANG stock, faceted by stock symbol.
+@fig-outcome-var illustrates the distribution of the adjusted closing prices (**'adjusted'**) for each FAANG stock, faceted by stock symbol.
 These distributions highlight the variability in stock prices across the companies, with stocks like NFLX showing higher price ranges, whereas GOOGL and AMZN have narrower distributions.
@@
 
-Together, these variables provide a robust framework for understanding and predicting stock price movements over time, leveraging both raw market indicators and derived metrics.
+Together, these variables provide a robust framework for understanding and predicting stock price movements over time, utilizing both raw market indicators and derived metrics.
 
@@
 set.seed(123) # For reproducibility
-library(dplyr)
-library(knitr)
 
@@
 All values are rounded to two decimal places to enhance clarity and precision.
-This table illustrates the diverse set of historical and technical indicators used to predict future adjusted stock prices.
 
@@
 
-The models in this study forecast the adjusted closing prices of FAANG stocks (Meta, Amazon, Apple, Netflix, and Google) for the months of November and December 2024.
-Separate models are constructed for each month to capture the time-dependent nature of stock price movements and to account for fluctuations in market conditions leading up to the forecast period.
-These models utilize a combination of historical stock price data and technical indicators, such as moving averages, volatility, and lagged price features, to estimate future stock prices for each of the five FAANG companies.
-
-To model the changes in stock prices over time, we use features that capture both short-term and long-term market trends.
-The models incorporate lagged stock prices, 7-day rolling averages, 20-day and 50-day simple moving averages (SMA), daily returns, and volatility indicators.
-These features are crucial for understanding the patterns and cycles in stock price movements.
-XGBoost, a powerful machine learning algorithm, is used to build these models, allowing for the capture of non-linear relationships and interactions between the technical indicators and future stock prices.
-
-The training datasets for the November and December models are defined using data up to October 31, 2024, and November 30, 2024, respectively.
-The models take into account temporal changes in stock prices by using features like lagged prices and rolling averages, which help capture the dynamics of stock price behavior.
-Additionally, XGBoost's flexibility allows the models to account for varying levels of volatility and market shifts across the different FAANG stocks.
-
-This modeling approach enables the prediction of future stock prices, considering how past price trends and market indicators evolve.
-By focusing on the November and December timeframes, the models aim to provide forecasts that reflect recent market conditions and predict likely price movements as we approach the end of 2024.
-The detailed model specifications, including feature selection and hyperparameter tuning, are further discussed in [Appendix -@sec-model-details]
+The XGBoost regression models in this study predict adjusted closing prices for FAANG stocks—Meta, Amazon, Apple, Netflix, and Google—for November and December 2024.
+Separate models are constructed for each month to account for time-specific market dynamics and recent trends.
+To capture short and long-term trends in stock price movements, these models incorporate features such as lagged prices, moving averages, daily returns, and volatility.
+Training data is segmented by timeframe, with the November model using data up to October 31, 2024, and the December model using data through November 30, 2024.
+Temporal features and technical indicators enable the models to capture stock-specific volatility and broader market trends.
+Further details on model specifications and diagnostics are provided in [Appendix -@sec-model-details].
 
@@
 
-To ensure the validity and reliability of our stock price prediction models, several key assumptions are made during the modeling process.
-These assumptions are necessary to capture the complexities of stock price behavior while maintaining a clear and interpretable framework for forecasting:
+To ensure the validity and reliability of our XGBoost models, several key assumptions are made during the modeling process:
 
@@
 We assume that its regularization techniques will help prevent overfitting while still capturing the intricate patterns present in the stock price data.
-
-These assumptions form the foundation of the models and guide the interpretation of the results.
-While they allow for effective forecasting, it is important to acknowledge that stock price prediction is inherently uncertain, and the accuracy of predictions may vary depending on unforeseen market events and fluctuations.
 
@@
 
-The primary goal of the November model is to predict the adjusted closing prices of FAANG stocks for the month of November 2024 based on historical stock price data.
-This model is structured as a regression task using the XGBoost algorithm to estimate the future adjusted closing prices for each of the five FAANG companies.
-
-The model setup can be described as follows:
+The primary estimand of the November model is the adjusted closing price of FAANG stocks for November 2024.
+The model is structured as a regression task using the XGBoost algorithm to estimate these prices for each of the five FAANG companies.
+The model setup is as follows:
 
@@
 
-The dataset for the November model is prepared by extracting the relevant data from January 2020 to October 31, 2024.
+The data for the November model is prepared by extracting the relevant data from January 2020 to October 31, 2024.
 Features like moving averages and volatility are calculated based on this historical data, and the model is trained to predict the adjusted closing prices for November 2024.
 
-The model was implemented using the **`xgboost`** package in R, which is optimized for performance in regression tasks:
+The model was implemented using the `xgboost` of @xgboost in R, which is optimized for performance in regression tasks:
 
@@
 $$
-
-Where $y_i$ represents the actual adjusted closing price for day $i$ and $\theta$ represents the model parameters to be learned during training. The model is trained with the following key parameters:
-- **Max Depth**: 6
-- **Learning Rate**: 0.1
-- **Number of Rounds**: 100
-- **Objective Function**: Regression with squared error
+Where $y_i$ represents the actual adjusted closing price for day $i$ and $\theta$ represents the model parameters to be learned during training.
+
+Key hyperparameters for the model are specificed as:
+
+- **Max Depth**: 6, to control the complexity of the trees and prevent overfitting
+- **Learning Rate**: 0.1, to balance the contribution of each boosting iteration
+- **Number of Rounds**: 100, to determine the total iterations for model optimization
+- **Objective Function**: Regression with squared error, suitable for continuous target variables.
+
+This set-up ensures that the model captures patterns in stock price movements while balancing bias and variance for improved predictive performance.
+Key assumptions include the appropriateness of historical stock price trends and technical indicators for predicting future prices.
 
@@
 
-The December model follows a similar approach to the November model, but is trained using data up to November 30, 2024.
-This ensures that the model accounts for the most recent stock price movements leading up to December, providing a more accurate prediction for the final month of 2024.
-
-The setup for the December model mirrors that of the November model, with the same features and target variable, but applied to the updated dataset.
+The December model follows a similar approach to the November model but incorporates extrapolated data for December 2024.
+Placeholder dates for December are added, and missing values in the adjusted price column are filled using forward and backward interpolation.
+This ensures the model reflects recent trends while accounting for the data cutoff on November 30, 2024.
 The model structure remains the same, and the XGBoost algorithm is again used for regression:
@@
 
-The December model is trained using the same XGBoost parameters as the November model.
+The December model is trained using the `XGBoost` of @xgboost library and the same parameters as the November model.
 Once trained, the model will generate predictions for December 2024, leveraging the latest market conditions reflected in the training data.
@@
 Additionally, the model's flexibility helps capture the intricacies of stock price behavior over time, ensuring that past market patterns are reflected in future predictions.
+Details are further discussed in [Appendix -@sec-model-justification]
 
@@
 
+## Model summary
+Both models were evaluated using RMSE, MAE, and R² metrics, as shown below in @fig-model-summary.
+The November model achieved an RMSE of 40.50, MAE of 16.40, and R² of 0.988, indicating it captured meaningful variability in the stock price movements.
+The December model reported significantly lower RMSE (1.10) and MAE (0.78), with an R² of 0.999, suggesting near-perfect alignment with the data.
+However, the December model's near-flat projections likely stem from minimal variability in the placeholder data used for December, resulting in artificially inflated performance metrics.
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-model-summary
+#| fig-cap: "November and December Model Summary"
+#| warning: false
+
+# Filter test data for November and December
+nov_test <- analysis_data %>%
+  filter(as.Date(date) > as.Date("2024-10-31") & as.Date(date) <= as.Date("2024-11-30"))
+
+dec_test <- analysis_data %>%
+  filter(as.Date(date) > as.Date("2024-11-30") & as.Date(date) <= as.Date("2024-12-31"))
+
+# Prepare test features and labels for November
+X_test_nov <- nov_test %>%
+  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+  as.matrix()
+y_test_nov <- as.numeric(nov_test$adjusted)
+
+# Prepare test features and labels for December
+X_test_dec <- dec_test %>%
+  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+  as.matrix()
+y_test_dec <- as.numeric(dec_test$adjusted)
+
+# Generate predictions
+nov_preds <- predict(nov_model, X_test_nov)
+dec_preds <- predict(dec_model, X_test_dec)
+
+# Calculate metrics for November
+nov_metrics <- tibble(
+  Metric = c("RMSE", "MAE", "R²"),
+  November = c(
+    rmse_vec(y_test_nov, nov_preds),
+    mae_vec(y_test_nov, nov_preds),
+    rsq_vec(y_test_nov, nov_preds)
+  )
+)
+
+# Calculate metrics for December
+dec_metrics <- tibble(
+  Metric = c("RMSE", "MAE", "R²"),
+  December = c(
+    rmse_vec(y_test_dec, dec_preds),
+    mae_vec(y_test_dec, dec_preds),
+    rsq_vec(y_test_dec, dec_preds)
+  )
+)
+
+# Combine metrics into one table
+combined_metrics <- nov_metrics %>%
+  left_join(dec_metrics, by = "Metric")
+
+# Display combined table using kable
+kable(combined_metrics)
+```
+These results highlight the November model's relative robustness in capturing dynamic trends compared to the December model's overfitted nature due to static input data.
+This emphasizes the importance of data variability in evaluating predictive model performance.
+
 # Results {#sec-results}
 
-Using the models trained on historical data, we applied the XGBoost algorithm to predict the adjusted closing prices for each of the FAANG stocks during these two months.
-For November, we filtered the data from November 1st to 30th, 2024, and prepared the relevant features for prediction, including lagged prices, rolling means, and volatility indicators.
-Since we have the actual adjusted closing prices for November, we compare the predicted values with the actual data to evaluate model performance.
-For December, the same approach was applied to the data from December 1st to 31st, 2024, but since the actual prices for this month are unavailable, the results for December focus solely on the predicted values.
+Using the models trained on historical data, we implemented the XGBoost algorithm to predict the adjusted closing prices for each of the FAANG stocks during these two months.
+For November, since we have the actual adjusted closing prices for November, we compare the predicted values with the actual data to evaluate model performance.
+For December, since the actual prices for this month are unavailable, the results for December focus solely on the predicted values.
 
@@
 #| label: fig-nov-actual-pred
-#| fig-cap: "Actual vs Predicted Stock Prices for November 2024"
+#| fig-cap: "Comparison of actual (blue) vs.
+predicted (red) adjusted closing stock prices for FAANG companies in November 2024.
+Highlights the model's ability to capture general trends but shows variability in performance across stocks, particularly for Netflix (NFLX)"
 #| warning: false
@@
 ```
-@fig-nov-actual-pred presents the comparison of actual versus predicted stock prices for the FAANG companies during November 2024.
-For each company—Apple (AAPL), Amazon (AMZN), Meta (META), Netflix (NFLX), and Google (GOOGL)—we plot the actual adjusted closing prices (in blue) alongside the predicted values (in red).
-The predictions, derived from our XGBoost models, demonstrate a generally strong alignment with the observed stock price movements, particularly for AAPL and META, where the predicted prices closely follow the actual trends.
-However, for companies like NFLX, the predicted values show a less consistent match, especially later in the month.
-This discrepancy suggests that while the model captures major trends effectively, there may be some volatility or market events that were not fully accounted for in the prediction process.
-These visualizations help assess the performance of our models and highlight areas where the predictions are robust as well as where there is room for improvement.
-
-## Heatmap of Prediction Errors (November 2024)
-```{r}
-#| echo: false
-#| eval: true
-#| label: fig-heatmap
-#| fig-cap: "Heatmap of Prediction Errors for November 2024"
-#| warning: false
-
-# Add the `error` column to `nov_results`
-nov_results <- nov_results %>%
-  mutate(error = abs(adjusted - predicted_adjusted))  # Calculate absolute error
-
-# Prepare data for heatmap
-nov_heatmap_data <- nov_results %>%
-  select(date, symbol, error) %>%
-  spread(symbol, error)
-
-# Plot heatmap of prediction errors
-ggplot(melt(nov_heatmap_data, id.vars = "date"), aes(x = date, y = variable, fill = value)) +
-  geom_tile() +
-  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
-  labs(
-    x = "Date",
-    y = "Stock Symbol",
-    fill = "Error"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    text = element_text(size = 12)
-  )
-```
-@fig-heatmap highlights the discrepancies between the predicted and actual stock prices for each FAANG company on a daily basis.
-The color intensity, ranging from light pink to deep red, indicates the magnitude of the errors, with darker shades representing higher errors.
-Notably, the Netflix (NFLX) stock shows consistently higher errors throughout November, particularly on November 4th and November 11th, where the error values reach above 100, as indicated by the deeper red color.
-Meta (META) and Google (GOOGL) exhibit moderate errors, with values peaking around November 11th but staying relatively consistent compared to NFLX.
-In contrast, Amazon (AMZN) and Apple (AAPL) show smaller error values, especially in the first half of the month, where the colors are much lighter.
-This heatmap provides a clear view of which stocks and dates the model struggled with the most, offering insights into potential areas for further model improvement.
+@fig-nov-actual-pred compares actual (blue) and predicted (red) stock prices for FAANG companies during November 2024.
+The model captures trends effectively for AAPL and META, where predictions align closely with observed prices.
+However, for NFLX, discrepancies appear later in the month, indicating the model's limitations in capturing certain market volatilities.
 
@@
 #| label: fig-nov-avg
-#| fig-cap: "Average Predicted Stock Prices for November 2024"
+#| fig-cap: "Average predicted stock prices for FAANG companies in November 2024.
+Netflix leads with the highest predicted average price (~$750), while Google has the lowest (~$180), reflecting varied performance across companies based on model predictions."
 #| warning: false
@@
 ```
-@fig-nov-avg illustrates the average predicted stock prices for each FAANG company in November 2024.
-The predictions reveal significant variations across the companies, with Netflix (NFLX) showing the highest predicted average price, approximately **$600**, followed by Meta (**'META'**) with an average just above **$500**.
-In contrast, Apple (**'AAPL'**), Amazon (**'AMZN'**), and Google (**'GOOGL'**) exhibit considerably lower average predicted prices, with AAPL at around **$225**, AMZN at **$210**, and GOOGL at approximately **$180**.
-This visualization highlights the distinct predicted performance of each company, suggesting that Netflix is expected to outperform the others in terms of stock price in November 2024, based on our model’s predictions.
-These results provide a clear comparative overview of expected stock movements and serve as a foundation for further analysis of market trends in the upcoming months.
+@fig-nov-avg shows the average predicted stock prices for FAANG companies in November 2024.
+Netflix (NFLX) has the highest predicted average price (~$750), followed by Meta (META) at just above $500.
+Apple (AAPL), Amazon (AMZN), and Google (GOOGL) exhibit lower averages, at ~$225, ~$210, and ~$180, respectively.
+The chart highlights Netflix's expected strong performance compared to other FAANG companies.
 
@@
 
-The line chart above presents the predicted stock prices for each FAANG company throughout December 2024.
-The predicted prices for each stock exhibit a strikingly consistent slope, with no significant fluctuations or adjustments throughout the month.
-For example, Netflix (NFLX) shows a steady increase, while Apple (AAPL), Amazon (AMZN), and Google (GOOGL) display relatively flat predicted price trends.
-This consistent pattern suggests potential limitations in the model, as real-world stock prices are typically subject to more volatility and market dynamics that are not captured by this prediction.
-
 ```{r}
@@
 #| label: fig-dec-pred
-#| fig-cap: "Predicted Stock Price Trends for December 2024"
+#| fig-cap: "Predicted stock prices for FAANG companies in December 2024.
+The model predicts relatively flat price trends for all companies, with Netflix having the highest predicted price and Google the lowest"
 #| warning: false
@@
 ```
-The lack of variability in the predictions may indicate that the model fails to account for potential market shifts, such as reactions to economic news, company announcements, or other external factors.
-Additionally, the absence of any sharp price movements could suggest that the model overly smooths the data, possibly due to the features used or the nature of the training data.
-These limitations highlight the importance of refining the model to incorporate broader market influences and improve its responsiveness to sudden changes in the stock market environment.
+@fig-dec-pred presents the predicted stock prices for FAANG companies in December 2024, showing flat trends with no significant fluctuations.
+Netflix (NFLX) has the highest predicted price, while Google (GOOGL) has the lowest.
+Such flat trends highlight potential limitations in the model's ability to capture real-world stock price volatility, further detailed in [Appendix -@sec-model-limitations].
 
@@
 #| label: fig-dec-avg
-#| fig-cap: "Average Predicted Stock Prices for December 2024"
+#| fig-cap: "Average predicted stock prices for FAANG companies in December 2024.
+Netflix leads with the highest predicted price ($825), followed by Meta ($600), while Apple, Amazon, and Google show considerably lower averages."
 #| warning: false
@@
 ```
-The bar chart above displays the average predicted stock prices for each FAANG company in December 2024.
-As shown, Netflix (NFLX) is expected to have the highest average predicted stock price, just above **$750**, followed by Meta (META) at approximately **$600**.
-In contrast, Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show significantly lower predicted prices, with AAPL around **$250**, AMZN slightly lower, and GOOGL at just above **$200**.
-This suggests that, according to the model, Netflix is projected to continue outperforming other FAANG companies in terms of stock price.
-These predictions reflect market trends modeled from historical data, with a noticeable discrepancy between the higher-performing stocks and the others.
-The chart provides a comparative view of how each stock is expected to fare during December 2024, contributing to our understanding of potential market movements for the month.
-
-## Average Predicted Stock Price - December
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-dec-avg
-#| tbl-cap: "Average Predicted Stock Prices for December 2024"
-#| warning: false
-
-# Calculate average predicted price for December
-dec_avg_prices <- dec_results %>%
-  group_by(symbol) %>%
-  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE)) %>%
-  mutate(avg_predicted_price = dollar(round(avg_predicted_price, 2))) %>%  # Format as dollar and round
-  arrange(desc(avg_predicted_price))  # Sort by average price in descending order
-
-# Display the table
-dec_avg_prices %>%
-  kable(
-    caption = "Average Predicted Stock Prices for December 2024",
-    col.names = c("Stock Symbol", "Average Predicted Price"),
-    format = "markdown"  # Output as markdown for readability
-  )
-```
-@tbl-dec-avg presents the average predicted stock prices for each FAANG company in December 2024.
-According to the predictions, Netflix (NFLX) is expected to have the highest average stock price at **$887.84**, followed by Meta (META) with an average of **$574.92**.
-Apple (AAPL) is predicted to have an average stock price of **$235.37**, while Amazon (AMZN) and Google (GOOGL) are expected to have lower average prices, at **$207.93** and **$168.69**, respectively.
-These predicted values highlight significant variations across the FAANG companies, with Netflix forecasted to substantially outperform the others.
-This table offers a clear numerical summary of the expected market performance for each company, providing insights into potential market trends for December 2024.
+@fig-dec-avg shows the average predicted December 2024 stock prices for FAANG companies.
+Netflix (NFLX) leads with the highest predicted price, just around $825, followed by Meta (META) at $600.
+Apple (AAPL), Amazon (AMZN), and Google (GOOGL) are projected significantly lower, around $250, $200, and $200, respectively.
+These predictions highlight Netflix's strong expected performance compared to its peers, reflecting trends modeled from historical data.
 
@@
 #| label: fig-nov-dec-pred
-#| fig-cap: "Predicted Stock Prices for November and December 2024"
+#| fig-cap: "Predicted stock prices for FAANG companies across November (red) and December (teal) 2024.
+November predictions exhibit variability, particularly for Apple (AAPL), Meta (META), Amazon (AMZN), and Google (GOOGL), while December predictions flatten, showing minimal fluctuations.
+Netflix (NFLX) remains nearly flat across both months.
+Highlights the model's reduced sensitivity to potential market dynamics in December."
 #| warning: false
@@
 ```
-@fig-nov-dec-pred above illustrates the predicted stock prices for each FAANG company across November and December 2024.
-The predictions are shown for both months, with the November predictions represented by the teal lines and the December predictions shown in red.
-For companies like Apple (AAPL) and Meta (META), the model predicts a notable increase in stock prices during November, followed by a flat trend throughout December, indicating minimal movement in the predicted prices during the second month.
-In contrast, for Netflix (NFLX), the predicted stock price remains almost flat through both months, showing limited variation.
-Amazon (AMZN) and Google (GOOGL) exhibit more fluctuations in November, with predicted prices stabilizing into flat trends for December.
-This pattern highlights a potential limitation of the model, as it fails to capture significant market dynamics or volatility in December, suggesting that the model may not fully account for market influences or external factors that could impact stock prices at the end of the year.
+@fig-nov-dec-pred illustrates predicted FAANG stock prices for November (teal) and December (red) 2024.
+November shows dynamic trends for Apple (AAPL), Meta (META), Amazon (AMZN), and Google (GOOGL), while December predictions flatten across all companies, particularly Netflix (NFLX).
+This highlights the model's limitation in capturing market volatility and end-of-year dynamics.
 
@@
 #| label: fig-nov-dec-avg
-#| fig-cap: "Average Predicted Stock Prices for November and December 2024"
+#| fig-cap: "Average predicted stock prices for FAANG companies in November (red) and December (teal) 2024.
+Netflix shows the highest increase between months, while Apple, Amazon, and Google exhibit stable predicted prices.
+Only google displays month-to-month decrease."
 #| warning: false
@@
 ```
-@fig-nov-dec-avg compares the average predicted stock prices for each FAANG company in November and December 2024.
-The chart clearly shows the predicted stock prices for November (red bars) and December (teal bars).
-Notably, Netflix (NFLX) is expected to experience the highest increase in stock price, with a marked difference between November and December.
-Meta (META) follows with a similar pattern, although the difference between months is less pronounced.
-In contrast, stocks like Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show relatively stable predicted prices between the two months, with only slight variations.
-These results suggest that while some stocks, like Netflix, are predicted to have larger fluctuations, others show minimal change in their predicted prices, indicating potential stability or lack of significant market events influencing these stocks during December 2024.
+@fig-nov-dec-avg compares the average predicted stock prices for FAANG companies in November (red) and December (teal) 2024.
+Netflix (NFLX) shows the highest predicted prices and a notable increase from November to December.
+Meta (META) also exhibits a slight increase, while Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show minimal variations, indicating relative stability in their predicted prices across the two months.
+Only Google (GOOGL) displays a slight month-to-month decrease, suggesting for further validation and data analysis.
 
@@
 
-@tbl-election-effect illustrates the stock prices of FAANG companies one trading day before and after the elections for the years 2016, 2020, and 2024.
-The post-election period often shows noticeable fluctuations in stock prices, with some stocks experiencing a sharp increase and others a decrease.
-For example, AAPL saw a slight increase in 2016 and 2020, while AMZN and NFLX both exhibited declines in 2016.
-In contrast, the 2024 data reveals an overall upward trend across all FAANG stocks, with particularly strong gains for NFLX, which rose from $755.51 before the election to $780.21 after the election, reflecting a surge in stock prices.
-
+@tbl-election-effect compares FAANG stock prices one trading day before and after the elections in 2016, 2020, and 2024.
+Post-election periods show noticeable fluctuations, with some stocks rising and others declining.
+For instance, AAPL experienced slight increases in 2016 and 2020, while AMZN and NFLX declined in 2016.
+In contrast, 2024 shows an overall upward trend across all FAANG stocks, with NFLX seeing the largest gain, rising from $755.51 to $780.21.
 ```{r}
@@
 ```
-This post-election stock behavior suggests that broader market reactions to election outcomes may have impacted stock prices significantly, particularly towards the end of the election year.
-Given this, it is likely that the increase in stock prices in late 2024 contributed to the predicted stock price trends in our model.
-As our model leverages historical data and trends starting from 2020, including fluctuations in stock prices towards the end of 2024, this surge likely influenced the final predictions, making them more optimistic and reflecting the strong market response post-election.
-This potential influence should be considered when interpreting the results, as it emphasizes the role of market sentiment and external factors that may cause sharp fluctuations in stock prices, beyond the intrinsic factors captured by the model.
+These trends suggest that market reactions to election outcomes significantly influenced stock prices, particularly in late 2024.
+This surge likely contributed to the increased stock price predictions for December in our model, which incorporates historical data and trends.
+The impact of market sentiment and external factors, such as post-election dynamics, highlights the need to consider broader influences when interpreting the model's results.
 
@@
 
-# Additional data details {#sec-data-details}
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-## Diagnostics
+# Additional Data Details {#sec-data-details}
+
+## Data Cleaning {#sec-data-cleaning}
+
+The data cleaning process transformed raw FAANG stock price data from Yahoo Finance into a structured format suitable for predictive modeling.
+The dataset spans daily stock prices for Meta (META), Amazon (AMZN), Apple (AAPL), Netflix (NFLX), and Alphabet (GOOGL) from January 2020 to November 2024.
+To prepare for December 2024 predictions, placeholder dates were added to create a continuous sequence of daily observations, accounting for non-trading days such as weekends and holidays.
+Missing values in the adjusted closing price column were addressed through forward and backward interpolation, ensuring data continuity for subsequent analyses.
+Rigorous testing ensured no critical missing values remain, all values fall within logical ranges, and the dataset maintain consistency.
+
+## Feature Engineerring {#sec-feature-engineering}
+Feature engineering enriched the dataset with additional predictors to improve model performance.
+Key features included Lag_1 (adjusted closing price from the previous day), Rolling_Mean_7 (7-day rolling average), sma_20 and sma_50 (20-day and 50-day simple moving averages), volatility (20-day rolling standard deviation), and daily_return (percentage change in adjusted closing price).
+These features captured temporal trends, smoothed short-term fluctuations, and quantified variability in stock prices.
+Stock symbols were numerically encoded to help the model identify patterns across different stocks.
+These features are further explained in [Appendix -@sec-feature-details]
+
+## Data Limitations and Future Directions {#sec-data-limitations}
+
+While the cleaned dataset is robust and suitable for modeling, several limitations remain.
+External factors, such as macroeconomic indicators, earnings announcements, or industry-specific news, are not included, despite their potential to significantly impact stock prices.
+The interpolation of missing values assumes smooth price continuity, which may not fully account for abrupt market changes or high-impact events occurring during non-trading days.
+Additionally, the analysis is restricted to FAANG stocks, which, although influential, may not capture broader market dynamics or indirect effects from other sectors.
+These limitations may reduce the model's ability to fully represent the complexities of the market.
+
+## Furutre Directions
+
+Future analyses could address these limitations by incorporating additional data sources, such as macroeconomic indicators or sector-specific variables, to enhance predictive accuracy.
+Expanding the dataset to include other major stocks or sectors could provide a more comprehensive view of market behavior and its impact on FAANG stocks.
+Advanced feature engineering techniques, such as seasonal decomposition, non-linear transformations, or incorporating time-varying covariates, could improve the models’ ability to capture complex interactions and nuanced market behaviors.
+These improvements would strengthen the robustness and generalizability of the analysis, offering deeper insights into both short- and long-term trends in FAANG stock dynamics.
+
+# Additional Model Details {#sec-model-details}
+
+## Model Justification - Further Details
+
+The XGBoost regression model was chosen for predicting FAANG stock prices in December 2024 due to its ability to model complex non-linear relationships, which are common in financial time series data.
+Given the dynamic and time-sensitive nature of stock prices, the November and December models were trained separately to ensure that predictions account for evolving market trends and feature importance within each specific period.
+
+### Why Separate November and December Models?
+
+Separate models for November and December were developed to capture the unique characteristics of stock price movements during these months.
+By training the November model using data up to October 31, 2024, and the December model using data up to November 30, 2024, we ensure that each model incorporates the most recent market information available for its respective forecasting period.
+This separation minimizes over-fitting to earlier trends and allows the December model to adapt to any shifts in market behavior or volatility observed in November.
+
+### Model Features {#sec-feature-details}
+
+XGBoost was selected due to its ability to handle non-linear interactions and temporal dependencies, which are crucial for stock price prediction. Key engineered features include:
+
+- **Lag_1**: Represents the adjusted closing price from the previous day, capturing short-term price dependency.
+- **Rolling_Mean_7**, **sma_20**, **sma_50**: Reflect short- and medium-term trends, offering insights into market behavior over recent periods.
+- **Volatility**: Captures price variability using the 20-day rolling standard deviation, a critical factor in understanding market risk.
+- **Daily_Return**: Quantifies day-to-day price changes, highlighting rate-of-change trends.
+- **Symbol_Encoded**: Encodes stock-specific identifiers numerically, enabling the model to generalize patterns across multiple FAANG stocks.
+
+By incorporating these features, the model is able to account for short-term fluctuations and medium-term market dynamics, which are critical for accurate stock price forecasting.
+
+### Hyperparameters and Regularization
+
+The November and December models share key hyperparameters optimized to balance complexity and generalizability:
+- **Max depth**: Set to 6, allowing the model to capture significant interactions without overfitting.
+- **Learning rate**: Set to 0.1, ensuring gradual refinement of predictions.
+- **Boosting rounds**: Fixed at 100 to provide sufficient iterations for model convergence.
+XGBoost's built-in regularization techniques mitigate overfitting, especially important given the high-dimensional engineered features and the presence of missing values due to holidays and weekends.
+
+Unlike traditional time-series models such as ARIMA, which assume stationarity and often require extensive preprocessing, XGBoost directly handles non-stationary data.
+This flexibility makes it an ideal choice for stock price forecasting, where trends and volatility can vary significantly over time.
+Furthermore, XGBoost automatically handles missing data, making it robust for real-world applications with gaps in trading days.
+The decision to train separate models for November and December, coupled with XGBoost's ability to handle complex non-linearities, ensures that the forecasts reflect both temporal dependencies and evolving market dynamics.
+By balancing model flexibility and interpretability, the November and December XGBoost regression models provide a robust and reliable framework for predicting FAANG stock prices, tailored to the unique characteristics of each time period.
+
+## Model Diagnostics {#sec-model-diagnostics}
+
+### Feature Importance
+@fig-nov-feature-importance for the November model reveals that Lag_1 (the previous day's adjusted closing price) is the most influential predictor, accounting for the majority of the model's decision-making.
+Rolling_Mean_7 (the 7-day rolling average) is the second most significant feature, albeit with a much smaller contribution.
+Other features, such as daily_return, sma_20, and volatility, show negligible impact.
+This indicates that the model primarily relies on short-term trends and immediate past values to predict stock prices, reflecting its focus on short-term forecasting.
+```{r}
+#| context: global
+#| echo: false
+#| eval: true
+#| label: fig-nov-feature-importance
+#| fig-cap: "November Feature Importance"
+#| warning: false
+#| cache: false
+
+# Ensure the November model and feature data exist
+if (!exists("nov_model")) {
+  nov_model <- readRDS("models/nov_model.rds") # Load the pre-trained November model
+}
+
+  # Recreate X_train_nov if not present in the environment
+  nov_train <- analysis_data %>%
+    filter(as.Date(date) <= as.Date("2024-10-31"))
+
+  X_train_nov <- nov_train %>%
+    select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+    as.matrix()
+
+# Generate feature importance matrix
+importance_matrix <- xgb.importance(
+  feature_names = colnames(X_train_nov),
+  model = nov_model
+)
+
+# Plot feature importance
+xgb.plot.importance(
+  importance_matrix,
+  rel_to_first = TRUE,
+  top_n = 10,
+  main = "Feature Importance for November Model"
+)
+```
+
+#### December Model
+@fig-dec-feature-importance reveals that for the December model, Lag_1 remains the dominant feature, followed by Rolling_Mean_7, which has a slightly larger contribution compared to the November model.
+The remaining features, including daily_return, sma_20, and volatility, exhibit minimal importance, similar to the November model.
+The heavy reliance on short-term features like Lag_1 and Rolling_Mean_7 underscores the model's emphasis on recent price trends for prediction, while broader, longer-term patterns play a negligible role in its forecasting approach.
+```{r}
+#| context: global
+#| echo: false
+#| eval: true
+#| label: fig-dec-feature-importance
+#| fig-cap: "Feature Importance for December Model"
+#| warning: false
+#| cache: false
+
+  dec_train <- analysis_data %>%
+    filter(as.Date(date) <= as.Date("2024-11-30"))
+
+  X_train_dec <- dec_train %>%
+    select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+    as.matrix()
+
+# Generate feature importance matrix for December model
+importance_matrix_dec <- xgb.importance(
+  feature_names = colnames(X_train_dec),
+  model = dec_model
+)
+
+# Plot feature importance for December model
+xgb.plot.importance(
+  importance_matrix_dec,
+  rel_to_first = TRUE,
+  top_n = 10,
+  main = "Feature Importance for December Model"
+)
+
+```
+
+### Residual Analysis
+#### November Model
+@fig-nov-residual-analysis evaluates the November model's performance by analyzing the residuals (errors) against the predicted (fitted) values.
+The residuals vs.
+fitted values plot for the November model shows that residuals are evenly distributed around zero for most fitted values, suggesting that the model captures the central trend of the data well.
+However, there are slight patterns in residuals at higher fitted values, indicating some systematic deviations that the model may not fully account for.
+The majority of residuals fall within a narrow range, suggesting good model accuracy overall, but some outliers indicate the potential influence of unmodeled factors or noise in the data.
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-nov-residual-analysis
+#| fig-cap: "Residuals vs.
+fitted values for the November model.
+The residuals are centered around zero, indicating accurate predictions overall, though some systematic deviations and outliers are observed at higher fitted values."
+#| warning: false
+
+# November Model Training
+
+# Train on data up to October 31, 2024
+nov_train <- analysis_data %>%
+  filter(as.Date(date) <= as.Date("2024-10-31"))
+
+# Prepare November training features (X_train) and labels (y_train)
+X_train_nov <- nov_train %>%
+  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+  as.matrix()
+y_train_nov <- as.numeric(nov_train$adjusted)
+
+residuals_nov <- y_train_nov - predict(nov_model, X_train_nov)
+ggplot(data.frame(Fitted = predict(nov_model, X_train_nov), Residuals = residuals_nov), aes(x = Fitted, y = Residuals)) +
+  geom_point(alpha = 0.5) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  labs(x = "Fitted Values", y = "Residuals") +
+  theme_minimal()
+
+```
+
+#### December Model
+@fig-dec-residual-analysis reveals a similar pattern to the November model, with residuals mostly centered around zero.
+However, the residuals exhibit slightly tighter clustering, reflecting the December model's reliance on flatter trends in the data.
+Outliers are still present, but their distribution is more uniform across fitted values.
+This tighter residual range suggests the December model captures the data's central trend but is less sensitive to short-term volatility, as reflected in the flat December predictions.
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-dec-residual-analysis
+#| fig-cap: "Residuals vs.
+fitted values for the December model.
+The residuals are more tightly clustered around zero, reflecting the model's reliance on flatter trends but with less sensitivity to short-term volatility."
+#| warning: false
+
+# Prepare December training features (X_train_dec) and labels (y_train_dec)
+dec_train <- analysis_data %>%
+  filter(as.Date(date) <= as.Date("2024-11-30"))
+
+X_train_dec <- dec_train %>%
+  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
+  as.matrix()
+y_train_dec <- as.numeric(dec_train$adjusted)
+
+# Calculate residuals for the December model
+residuals_dec <- y_train_dec - predict(dec_model, X_train_dec)
+
+# Create Residuals vs. Fitted plot for December model
+ggplot(data.frame(Fitted = predict(dec_model, X_train_dec), Residuals = residuals_dec), aes(x = Fitted, y = Residuals)) +
+  geom_point(alpha = 0.5) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  labs(
+    x = "Fitted Values",
+    y = "Residuals"
+  ) +
+  theme_minimal()
+
+```
+
+### Error Distribution
+#### November Model
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-nov-error-distrib
+#| fig-cap: "Histogram of residuals for the November model.
+Residuals are centered around zero with most values falling within -2.5 to 2.5, reflecting accurate predictions with minimal systematic error."
+#| warning: false
+
+# Calculate residuals for the November model
+residuals_nov <- y_train_nov - predict(nov_model, X_train_nov)
+
+# Create a histogram for residuals
+ggplot(data.frame(Residuals = residuals_nov), aes(x = Residuals)) +
+  geom_histogram(binwidth = 0.1, color = "black", fill = "blue", alpha = 0.7) +
+  labs(
+    title = "Error Distribution for November Model",
+    x = "Residuals",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+```
+@fig-nov-error-distrib reveals that error distribution for the November model is approximately normal, with residuals centered around zero and a slight skewness at the tails.
+Most residuals are concentrated within the range of -2.5 to 2.5, indicating the model predicts stock prices with a high degree of accuracy.
+However, the presence of a few outliers suggests the model struggles with certain extreme price movements.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-dec-error-distribution
+#| fig-cap: "Histogram of residuals for the December model. Residuals are tightly centered around zero, indicating consistent predictions but limited sensitivity to abrupt stock price changes."
+#| warning: false
+#| cache: false
+
+# Calculate residuals for the December model
+residuals_dec <- y_train_dec - predict(dec_model, X_train_dec)
+
+# Create a histogram for residuals
+ggplot(data.frame(Residuals = residuals_dec), aes(x = Residuals)) +
+  geom_histogram(binwidth = 0.1, color = "black", fill = "blue", alpha = 0.7) +
+  labs(
+    title = "Error Distribution for December Model",
+    x = "Residuals",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+```
+@fig-dec-error-distrib shows an approximate normal error distribution for the December Model, with residuals tightly clustered around zero.
+Compared to the November model, the distribution is slightly narrower, suggesting reduced variability in prediction errors.
+This reflects the December model's reliance on flatter trends and its inability to capture sharp price movements, leading to smaller but more consistent errors.
+
+### Prediction vs. Actual Values (November)
+@fig-model-accuracy demonstrates a near-perfect alignment between predicted and actual stock prices, as indicated by the data points closely following the 45-degree reference line.
+This suggests that the model accurately captures the stock price trends with minimal deviation.
+The tight clustering of points along the line reflects the model's strong predictive performance, supported by the high R² value observed in the evaluation metrics.
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-model-accuracy
+#| fig-cap: "Predicted vs. actual stock prices for the November model. The points closely follow the 45-degree line, indicating strong predictive accuracy and alignment with actual values."
+#| warning: false
+
+ggplot(data.frame(Predicted = predict(nov_model, X_train_nov), Actual = y_train_nov), aes(x = Predicted, y = Actual)) +
+  geom_point(alpha = 0.5) +
+  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
+  labs(title = "Predicted vs Actual Values", x = "Predicted Values", y = "Actual Values") +
+  theme_minimal()
+
+```
+Such visualization cannot be replicated for the December model because the predicted values for December are based entirely on placeholder dates, with no actual stock price data available for comparison.
+As such, there are no actual values for December against which to evaluate the predictions, making it impossible to construct a meaningful predicted vs.
+actual plot.
+
+## Model Performance and Limitations
+
+### November Model Performance
+The November model demonstrates strong predictive accuracy, as shown by the close alignment between predicted and actual values along the 45-degree reference line.
+Residuals are centered around zero and follow an approximately normal distribution, with most residuals falling within the range of -2.5 to 2.5.
+These results indicate that the model effectively captures short-term trends and variability in FAANG stock prices.
+Feature importance analysis highlights Lag_1 (previous day’s adjusted price) and Rolling_Mean_7 (7-day rolling average) as the most influential predictors, confirming the model's ability to leverage recent price dynamics for accurate forecasting.
+However, the model exhibits limitations in capturing extreme price movements or unusual events not reflected in the historical data.
+Slight patterns in residuals at higher fitted values and outliers in the error distribution suggest that the model struggles to account for high-impact or sector-specific factors, limiting its generalizability to unexpected market conditions.
+
+### December Model Performance
+The December model, trained on data up to November 30, 2024, shows significant limitations in its predictive capabilities.
+Residuals are tightly clustered around zero, with a narrower error distribution compared to the November model.
+While this suggests minimal noise in predictions, it reflects an overreliance on flat trends, resulting in an inability to capture dynamic stock price movements.
+The flat slopes in predicted values, combined with the absence of actual December stock price data for evaluation, underscore the model’s failure to incorporate potential market shifts, such as post-election sentiment or end-of-year trading dynamics.
+Similar to the November model, feature importance analysis identifies Lag_1 and Rolling_Mean_7 as dominant predictors.
+However, the December model’s heavy dependence on these features exacerbates its inability to account for broader market influences, leading to overly static predictions.
+
+### Model Limitations
+While the November and December XGBoost models demonstrate strong performance, several limitations should be noted:
+
+1.
+**Overreliance on Short-Term Features**: Both models heavily depend on Lag_1 and Rolling_Mean_7, which focus on short-term trends.
+While effective for stable markets, this reliance limits their ability to predict abrupt changes or long-term patterns influenced by broader macroeconomic or sector-specific factors.
+
+2.
+**Exclusion of External Influences**: Neither model incorporates external variables such as macroeconomic indicators, earnings reports, geopolitical events, or industry-specific news, which are critical in driving stock price movements.
+This omission reduces the models' capacity to adapt to dynamic market conditions.
+
+3.
+**Flat Predictions for Certain Periods**: The flat predictions observed in the December model highlight a broader limitation: both models assume continuity in trends and fail to account for potential volatility or sudden market shocks.
+This may be a result of insufficient diversity in the feature set.
+
+4.
+**Lack of Cross-Validation Across Time Periods**: While each model is trained separately on its respective time frame, there is no mechanism to ensure their consistency across overlapping or adjacent periods.
+This could lead to discrepancies when transitioning from November to December predictions.
+
+5.
+**Sensitivity to Data Availability**: The models rely entirely on historical stock price data, meaning any gaps or interpolation inaccuracies could propagate into the predictions.
+This limitation is particularly relevant for the December model, where placeholder dates and interpolated values were used for prediction.
+
+Addressing these limitations in future analysis, such as incorporating external data and improving the feature set to capture dynamic market behaviors, would enhance the robustness and applicability of the models.
+
+# Surveys, Sampling, and Observational Data
+
+## Nature of Observational Data
+
+The stock price data used in this study is observational, derived from naturally occurring financial transactions in the stock market.
+Unlike experimental or survey data, observational data reflects real-world behaviors influenced by external factors such as macroeconomic policies, geopolitical events, and investor sentiment.
+This complexity makes it challenging to distinguish causality from correlation, as unmeasured confounding variables can bias interpretations.
+For example, stock price movements during the 2024 U.S.
+Presidential Election may reflect both direct market reactions and broader economic conditions.
+
+While observational data offers a granular, high-frequency view of stock market dynamics, it relies entirely on historical trends as proxies for predictive features.
+This approach limits its ability to account for unprecedented shocks, such as global crises or election-related volatility, which may diverge from historical patterns.
+Furthermore, temporal dependence—where current prices are influenced by past values—requires careful feature engineering, such as lagged variables (Lag_1) and rolling averages (Rolling_Mean_7), to capture meaningful trends.
+
+## Sampling Frame and Representativeness
+The sampling frame encompasses all trading days from January 2020 to November 2024, providing a comprehensive record of FAANG stock activity during a period of significant market and political events.
+While this dataset ensures detailed coverage of the technology sector, it is constrained by its sector-specific focus, which limits generalizability to other industries or broader market trends.
+
+### Strengths and Weaknesses
+
+A key strength of the dataset is its high-frequency nature, which allows the analysis to capture both short-term fluctuations and long-term trends.
+FAANG stocks, as market leaders, serve as valuable indicators of overall economic conditions and investor sentiment.
+However, the focus on a single sector introduces limitations in generalizability, as trends within technology stocks may not reflect behaviors across other industries.
+Furthermore, the imputation of placeholder prices for December 2024 ensures continuity but risks underestimating volatility, particularly in the context of post-election market adjustments.
+
+## Integrating Surveys with Observational Data
+
+Surveys provide forward-looking indicators that enhance the interpretation of stock price trends.
+For example, during significant events like the 2024 U.S.
+Presidential Election, surveys could capture investor sentiment regarding anticipated market impacts, which may not yet be reflected in observed price movements.
+
+### Proposed Survey Design
+
+To complement the observational stock data, investor sentiment surveys could include questions such as:
+
+- **Market Confidence**: “Do you expect FAANG stock prices to increase, decrease, or remain stable over the next month?”
+- **Election Impact**: “How has the 2024 U.S. Presidential Election influenced your investment decisions?”
+- **Risk Appetite**: “What is your likelihood of investing in technology stocks in the current market climate?”
+
+### Methodological Considerations:
+
+- **Sampling and Representativeness**:  Stratify respondents by demographics, investment experience, and portfolio size to ensure diverse perspectives.
+- **Weighting**: Apply post-survey weighting to align survey demographics with actual investor populations, similar to post-stratification in traditional surveys.
+
+### Strengths and Limitations
+
+Advantages of Survey Integration:
+
+- **Contextualization**: Surveys can help interpret stock price trends, particularly around major events like elections.
+- **Predictive Power**: Sentiment data may improve model accuracy by capturing market expectations and behavioral drivers.
+
+Limitations of Survey Integration:
+
+- **Response Bias**: Survey participants may exaggerate optimism or pessimism, especially during volatile periods.
+- **Sampling Complexity**: Ensuring representativeness across demographics, such as institutional vs. retail investors, requires significant resources and methodological rigor
+
+## Parallels Between Observational Data and Surveys
+
+Observational stock data and surveys share structural similarities that make their integration intuitive and mutually reinforcing:
+
+- **Stratification**: The division of stock data by symbol mirrors stratified sampling in surveys, allowing for detailed analysis at the stock level while preserving aggregate insights.
+- **Imputation**: Both surveys and stock data require robust imputation methods to address missing values.
+For example, placeholder stock prices for December 2024 were interpolated using forward and backward filling, similar to imputing missing survey responses.
+- **Weighting**: Models could incorporate macroeconomic indicators, such as GDP or interest rates, as weights to adjust for external market influences, akin to post-stratification in surveys.
+
+## Addressing Data Challenges
+
+- **Temporal Dependence**: Stock prices are influenced by past trends.
+To mitigate this, features such as lagged prices (Lag_1) and rolling averages (Rolling_Mean_7) were used to capture momentum and smoothed trends effectively.
+However, these features may miss sudden shifts caused by external shocks or policy changes.
+
+- **Imputation for Missing Data**: Missing December 2024 values were filled through interpolation to ensure continuity.
+While necessary for analysis, this approach may underestimate volatility or fail to account for post-election market adjustments.
+Advanced time-series models could improve the accuracy of such imputations.
+
+- **Observational Bias**: Stock data is inherently influenced by external confounders, such as Federal Reserve policies, geopolitical tensions, or major corporate announcements.
+These confounders obscure causal relationships.
+Future models could include external predictors to better account for such influences, improving robustness.
+
+## Recommendations for Future Work
+
+1.
+**Survey Integration**: Combine investor sentiment surveys with stock data to capture qualitative dimensions of market behavior.
+Surveys can reveal forward-looking insights, such as investor confidence, that improve interpretability and predictive power.
+2. **Expanding Scope**: Extend the analysis to include sectors beyond FAANG stocks to enhance generalizability and uncover trends across diverse industries.
+3. **Improved Imputation**: Use dynamic time-series models to address missing values more accurately and capture volatility during critical events like elections or earnings announcements.
+
+By addressing these challenges and integrating surveys with observational data, future analyses can achieve a richer, more nuanced understanding of market dynamics.
+Such hybrid approach provides a comprehensive framework for interpreting stock price trends, making predictions more robust and actionable, especially during periods of significant economic or political change.
 

### 2024-12-04 paper/paper.qmd
committing updated paper.qmd and paper.pdf

@@
 library(yardstick)
+library(tidyquant)
+library(knitr)
 
@@
 #| label: fig-outcome-var
-#| fig-cap: "Distribution of Adjusted Closing Prices by Stock"
+#| fig-cap: "Distribution of Adjusted Closing Prices for FAANG Stocks (January 2020–November 2024).
+This visualization presents histograms of the adjusted closing prices for Apple (AAPL), Amazon (AMZN), Google (GOOGL), Meta (META), and Netflix (NFLX).
+Each panel illustrates the frequency of daily closing prices within specific price ranges, highlighting the variability in stock performance across the FAANG group during the observed period."
 #| warning: false
@@
 ```
-@fig-outcome-var illustrates the distribution of the adjusted closing prices (**'adjusted'**) for each FAANG stock, faceted by stock symbol.
-These distributions highlight the variability in stock prices across the companies, with stocks like NFLX showing higher price ranges, whereas GOOGL and AMZN have narrower distributions.
-The visual emphasizes the diverse price behavior among FAANG stocks, underscoring the need for tailored models that capture these unique trends.
-By normalizing for corporate actions like stock splits or dividends, the adjusted closing price serves as a reliable target variable for forecasting and analysis.
+@fig-outcome-var depict the distribution of adjusted closing prices for FAANG stocks over the period from January 2020 to November 2024.
+AAPL, AMZN, and GOOGL exhibit relatively compact distributions, with prices concentrated between $80–$200 for GOOGL and AMZN and $100–$250 for AAPL.
+META and NFLX show broader price ranges, with META peaking between $300–$400 and NFLX exhibiting the highest spread, spanning $250–$750.
+These distributions reflect the differing volatility and market dynamics of each stock, with META and NFLX experiencing higher fluctuations, likely influenced by company-specific events or broader market trends.
 
@@
 #| label: tbl-pred-var
-#| tbl-cap: "Random Sample of Predictive Variables for Modeling Adjusted Closing Price"
+#| tbl-cap: "Random Sample of Predictive Variables for Modeling Adjusted Closing Price.
+Showcases a sample of the predictive variables derived from raw stock data, including lagged prices, rolling averages, simple moving averages (SMA), volatility, daily returns, and symbol encodings."
 #| warning: false
@@
 ```
-@tbl-pred-var displays a random selection of five rows from the dataset, showcasing the predictive variables utilized in the modeling process.
-Each row represents data for a specific stock, including values for lagged adjusted closing prices (Lag_1), smoothed weekly averages (Rolling Mean (7-Day)), and medium- and long-term moving averages (SMA (20-Day) and SMA (50-Day)).
-These variables are complemented by measures of price volatility, daily percentage returns (Daily Return), and an encoded numerical identifier for the stock symbol (Symbol (Encoded)).
-All values are rounded to two decimal places to enhance clarity and precision.
+@tbl-pred-var provides a snapshot of the engineered predictive variables used in the modeling process.
+**'Lag_1'** reflects the adjusted closing price from the prior day, serving as a baseline for trend predictions.
+Rolling averages and simple moving averages (e.g., 7-day, 20-day, and 50-day) smooth out short-term fluctuations and emphasize broader price trends.
+Volatility, calculated as the 20-day rolling standard deviation, quantifies market uncertainty, while daily returns measure the day-to-day percentage change in adjusted closing price.
+The inclusion of symbol encodings ensures the model distinguishes between stocks while analyzing aggregate trends.
+These variables were carefully selected to balance short-term variability with medium-term patterns, enabling robust and reliable predictions.
 
@@
 predicted (red) adjusted closing stock prices for FAANG companies in November 2024.
-Highlights the model's ability to capture general trends but shows variability in performance across stocks, particularly for Netflix (NFLX)"
+Evaluates the November model’s predictive performance across Apple (AAPL), Amazon (AMZN), Google (GOOGL), Meta (META), and Netflix (NFLX)."
 #| warning: false
@@
 ```
-@fig-nov-actual-pred compares actual (blue) and predicted (red) stock prices for FAANG companies during November 2024.
-The model captures trends effectively for AAPL and META, where predictions align closely with observed prices.
-However, for NFLX, discrepancies appear later in the month, indicating the model's limitations in capturing certain market volatilities.
+@fig-nov-actual-pred indicates strong predictive performance for AAPL and META, where predicted prices closely follow actual trends, indicating the model's effectiveness in capturing both upward and downward market movements.
+AMZN and GOOGL exhibit moderate alignment, with some deviations occurring during periods of higher volatility, suggesting that while the model captures general trends, it struggles with abrupt changes.
+In contrast, NFLX demonstrates notable discrepancies, especially in the second half of November, where the predicted values fail to reflect the sharp upward trend observed in actual prices.
+These results highlight the model's capacity to replicate stable patterns but expose its limitations in adapting to rapid price movements, particularly for stocks like NFLX that experience heightened volatility.
 
@@
 ```
-@fig-nov-avg shows the average predicted stock prices for FAANG companies in November 2024.
-Netflix (NFLX) has the highest predicted average price (~$750), followed by Meta (META) at just above $500.
-Apple (AAPL), Amazon (AMZN), and Google (GOOGL) exhibit lower averages, at ~$225, ~$210, and ~$180, respectively.
-The chart highlights Netflix's expected strong performance compared to other FAANG companies.
+@fig-nov-avg displays that the November model predicts that Netflix (NFLX) will have the highest average stock price at approximately $750, reflecting its anticipated dominance among FAANG stocks.
+Meta (META) follows with an average price just above $500, while Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show significantly lower predicted averages, at around $225, $210, and $180, respectively.
+This result emphasizes the expected outperformance of Netflix and Meta, suggesting stronger market confidence in these two companies for November.
+Meanwhile, the comparatively lower averages for AAPL, AMZN, and GOOGL may indicate either more stable price trends or less aggressive growth potential modeled for these stocks during this period.
 
@@
 #| fig-cap: "Predicted stock prices for FAANG companies in December 2024.
-The model predicts relatively flat price trends for all companies, with Netflix having the highest predicted price and Google the lowest"
+Illustrates the predicted stock prices for FAANG companies during December 2024, highlighting flat trends across all stocks and minimal variation over the month."
 #| warning: false
@@
 ```
-@fig-dec-pred presents the predicted stock prices for FAANG companies in December 2024, showing flat trends with no significant fluctuations.
-Netflix (NFLX) has the highest predicted price, while Google (GOOGL) has the lowest.
-Such flat trends highlight potential limitations in the model's ability to capture real-world stock price volatility, further detailed in [Appendix -@sec-model-limitations].
+@fig-dec-pred reveals flat stock price trends for all FAANG companies, with no significant fluctuations throughout the month.
+Netflix (NFLX) is projected to maintain the highest price, exceeding $875, while Google (GOOGL) consistently exhibits the lowest price, just above $200.
+Apple (AAPL), Amazon (AMZN), and Meta (META) display similar stability, with their prices clustered between $250 and $600.
+These flat trends suggest that the model struggles to capture potential market dynamics and daily volatilities, likely due to the absence of real-world influences such as macroeconomic shifts or investor sentiment in the input data.
+Further discussion on these limitations is provided in [Appendix -@sec-model-limitations].
 
@@
 ```
-@fig-dec-avg shows the average predicted December 2024 stock prices for FAANG companies.
-Netflix (NFLX) leads with the highest predicted price, just around $825, followed by Meta (META) at $600.
-Apple (AAPL), Amazon (AMZN), and Google (GOOGL) are projected significantly lower, around $250, $200, and $200, respectively.
-These predictions highlight Netflix's strong expected performance compared to its peers, reflecting trends modeled from historical data.
+@fig-dec-avg shows that Netflix (NFLX) is expected to outperform its peers with an average price nearing $825, followed by Meta (META) at approximately $600.
+Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show significantly lower average predicted prices, clustering around $250, $200, and $200, respectively.
+These results suggest Netflix’s dominance in price performance among FAANG stocks, likely reflecting historical trends captured by the model.
+However, the flat trends and low variability in predictions suggest limitations in the model's capacity to incorporate dynamic market factors and real-world volatility.
 
@@
 November predictions exhibit variability, particularly for Apple (AAPL), Meta (META), Amazon (AMZN), and Google (GOOGL), while December predictions flatten, showing minimal fluctuations.
-Netflix (NFLX) remains nearly flat across both months.
-Highlights the model's reduced sensitivity to potential market dynamics in December."
+Netflix (NFLX) remains nearly flat across both months."
 #| warning: false
@@
 ```
-@fig-nov-dec-pred illustrates predicted FAANG stock prices for November (teal) and December (red) 2024.
-November shows dynamic trends for Apple (AAPL), Meta (META), Amazon (AMZN), and Google (GOOGL), while December predictions flatten across all companies, particularly Netflix (NFLX).
-This highlights the model's limitation in capturing market volatility and end-of-year dynamics.
+@fig-nov-dec-pred displays contrasting trends - November predictions reflect the market's dynamic nature, with AAPL and META showing upward trends and AMZN and GOOGL demonstrating fluctuations.
+Conversely, December predictions exhibit flat trends across all FAANG companies, particularly for NFLX, which remains static throughout the month.
+This stark contrast underscores the model's inability to adapt to market changes and capture potential end-of-year volatilities, limiting its effectiveness for December forecasting.
 
@@
 ```
-@fig-nov-dec-avg compares the average predicted stock prices for FAANG companies in November (red) and December (teal) 2024.
-Netflix (NFLX) shows the highest predicted prices and a notable increase from November to December.
-Meta (META) also exhibits a slight increase, while Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show minimal variations, indicating relative stability in their predicted prices across the two months.
-Only Google (GOOGL) displays a slight month-to-month decrease, suggesting for further validation and data analysis.
+@fig-nov-dec-avg compares the average predicted stock prices for FAANG companies in November (red) and December (teal) 2024, revealing varying dynamics across the companies.
+Netflix (NFLX) consistently leads with the highest prices, increasing significantly from November to December.
+Meta (META) follows a similar upward trend, albeit less pronounced.
+In contrast, Apple (AAPL), Amazon (AMZN), and Google (GOOGL) remain relatively stable, reflecting the model's ability to capture consistent trends for these stocks.
+Google's slight decrease between the months may suggest unique market dynamics requiring deeper investigation, particularly in the context of broader market sentiment and sector-specific factors.
 
@@
 #| label: tbl-election-effect
-#| tbl-cap: "FAANG Stock Prices for One Trading Day Before and After the Election (2016, 2020, and 2024)"
-#| warning: false
-
-library(tidyquant)
-library(dplyr)
-library(tidyr)
-library(knitr)
+#| tbl-cap: "Stock prices of FAANG companies one trading day before and after the U.S.
+Presidential Elections in 2016, 2020, and 2024.
+Most companies show an upward price change after the elections, particularly in 2024, where all FAANG stocks experienced increases, with Netflix (NFLX) recording the largest absolute gain."
+#| warning: false
 
@@
 #| label: fig-nov-feature-importance
-#| fig-cap: "November Feature Importance"
+#| fig-cap: "Feature importance for the November model.
+Lagged stock price (Lag_1) is the most significant predictor, followed by the 7-day rolling mean (Rolling_Mean_7).
+Other features, including daily returns, simple moving averages (sma_20, sma_50), volatility, and symbol encoding, contribute minimally to the model's predictions."
 #| warning: false
@@
 #| label: fig-dec-feature-importance
-#| fig-cap: "Feature Importance for December Model"
+#| fig-cap: "Feature importance for the December model.
+Highlights Lag_1 as the most influential predictor, followed by the 7-day rolling mean (Rolling_Mean_7).
+Other features, including daily returns, simple moving averages (sma_20, sma_50), volatility, and symbol encoding, have minimal contributions to the model's predictive capability."
 #| warning: false
@@
 fitted values for the November model.
-The residuals are centered around zero, indicating accurate predictions overall, though some systematic deviations and outliers are observed at higher fitted values."
+Residuals are centered around zero, indicating no major bias, but increased variability at higher fitted values suggests reduced accuracy for extreme stock prices."
 #| warning: false

### 2024-12-04 paper/paper.qmd
commit changes to preamble in R scripts

@@
 These companies represent diverse sectors, from e-commerce and technology to entertainment and digital advertising, making their stock price movements critical indicators of market health.
-In the context of the 2024 post-pandemic recovery and a pivotal United States Presidential Election, understanding and predicting FAANG stock price trends is particularly relevant for investors and policymakers alike.
-
-This paper focuses on historical stock price data from years 2020 to 2024, sourced from Yahoo Finance [@yahoofinance] to develop predictive models for FAANG stock prices in November and December 2024.
-Utilizing XGBoost regression models, we analyze historical trends and engineered features, including lagged prices, moving averages, and volatility measures, to forecast stock price movements.
-These predictions provide valuable insights into the behavior of FAANG stocks during critical periods, such as post-election market adjustments, and highlight opportunities and risks for short-term and long-term investment strategies.
+As the world navigates the post-pandemic recovery and the uncertainties surrounding the 2024 United States Presidential Election, understanding and forecasting FAANG stock prices is both timely and consequential for investors and policymakers alike.
+
+This paper examines FAANG stock price trends using historical data spanning 2020 to 2024, sourced from Yahoo Finance [@yahoofinance], to develop predictive models for November and December 2024.
+Utilizing XGBoost regression, the analysis leverages engineered features, including lagged prices, moving averages, and volatility measures, to predict adjusted closing prices for these stocks during the study period.
+By capturing trends and identifying potential risks and opportunities, these models provide actionable insights for short- and long-term investment strategies.
 
@@
 The implications of these findings extend beyond price predictions, offering a framework to evaluate market behavior in response to external factors such as elections and economic conditions.
-The remainder of this paper is organized as follows: Section 2 describes the data sources, variables, and preprocessing steps; Section 3 outlines the modeling approach and feature engineering strategies; Section 4 presents the results and discusses trends in predicted stock prices; and Section 5 addresses the implications and limitations of the findings, with additional model diagnostics detailed in the [Appendix -@sec-model-details].
+By contextualizing these trends within the broader market environment, this analysis contributes to a deeper understanding of market dynamics during critical periods.
+The remainder of this paper is organized as follows: @sec-data describes the data sources, variables, and preprocessing steps; @sec-model outlines the modeling approach and feature engineering strategies; @sec-results presents the results and discusses trends in predicted stock prices; and @sec-discussion addresses the implications and limitations of the findings, with additional model diagnostics detailed in the [Appendix -@sec-model-details].
 
@@
 These flat trends suggest that the model struggles to capture potential market dynamics and daily volatilities, likely due to the absence of real-world influences such as macroeconomic shifts or investor sentiment in the input data.
-Further discussion on these limitations is provided in [Appendix -@sec-model-limitations].
+Further discussion on these limitations is provided in [Appendix -@sec-model-perform-limit].
 
@@
 
-## Model Justification - Further Details
+## Model Justification - Further Details {#sec-model-justification}
 
@@
 
-## Model Performance and Limitations
+## Model Performance and Limitations {#sec-model-perform-limit}
 

### 2024-12-04 scripts/00-simulate_data.R

@@ -1,17 +1,15 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of FAANG stock prices
+# Author: Jamie Lee
+# Date: 30 November 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Any other information needed? N/A
 
 
 # Libraries
-library(dplyr)
-library(lubridate)
+library(tidyverse)
 
 # Set random seed for reproducibility
 set.seed(42)

### 2024-12-04 scripts/01-test_simulated_data.R

@@ -1,14 +1,13 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests simulated_data.csv
+# Author: Jamie Lee
+# Date: 30 November 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
+  # - 'testthat' package must be loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Any other information needed? N/A
 
 # Libraries
 library(testthat)

### 2024-12-04 scripts/02-download_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Yahoo Finance API
+# Author: Jamie Lee
+# Date: 30 November 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: tidyquant R package
+# Any other information needed? N/A
 
 
 # Load libraries

### 2024-12-04 scripts/03-clean_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans and prepares the raw FAANG stock data obtained from Yahoo Finance for predictive modeling
+# Author: Jamie Lee
+# Date: 30 November 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: raw_data.csv
+# Any other information needed? N/A
 
 # Libraries
 library(dplyr)

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests data quality of analysis_data.parquet
+# Author: Jamie Lee
+# Date: 4 December 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: load 'testthat' and analysis_data.parquet
+# Any other information needed? N/A
 
 # Libraries
 library(testthat)

### 2024-12-04 scripts/06-model_data.R

@@ -1,12 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Constructs november and december models for predicting stock prices using xgboost
+# Author: Jamie Lee
+# Date: 2 December 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: analysis_data.parquet
+# Any other information needed? N/A
 
 library(xgboost)
 library(dplyr)

### 2024-12-04 paper/paper.qmd
commit cross reference changes to paper.qmd
committing llm_usage.txt and survey.pdf; updating idealized methodology

@@
 
-The XGBoost regression models in this study predict adjusted closing prices for FAANG stocks—Meta, Amazon, Apple, Netflix, and Google—for November and December 2024.
+The XGBoost regression models in this study predict adjusted closing prices for FAANG stocks (Meta, Amazon, Apple, Netflix, and Google) for November and December 2024.
 Separate models are constructed for each month to account for time-specific market dynamics and recent trends.
@@
 
-## US Election Effects
+## Real-World Implications
+
+The predictive insights from this analysis hold significant value for investors, policymakers, and analysts, as FAANG stocks (Meta, Amazon, Apple, Netflix, Google) serve as key indicators of economic health and technological trends.
+Accurate stock price predictions enable informed decision-making in several areas:
+
+1.
+**Investment Strategies**: Anticipating price trends helps investors optimize portfolios.
+For instance, the predicted December 2024 price stability may guide conservative allocations or sector diversification.
+
+2. **Economic Forecasting**: FAANG performance offers insights into consumer behavior, innovation cycles, and macroeconomic conditions, aiding policymakers in evaluating sectoral resilience.
+
+3. **Risk Management**: Understanding volatility supports effective hedging strategies during high-risk periods, such as earnings reports or economic announcements.
+
+4.
+**Tech Sector Analysis**: Predictions inform strategic decisions for tech-driven businesses, highlighting market expectations for ad-reliant firms like Meta and Google or streaming services like Netflix.
+
+5. **Behavioral Insights**: Stock price trends reflect investor sentiment and external factors, emphasizing the need to integrate qualitative drivers, like trust in technology, into market models.
+
+These applications demonstrate the practical utility of FAANG stock predictions in navigating dynamic market environments and making data-driven decisions.
+
+### US Election Effect
+A critical component of stock market behavior is its sensitivity to major political events, such as U.S.
+presidential elections.
+These events often drive short-term volatility and influence investor sentiment, leading to measurable shifts in stock prices across sectors.
+Analyzing the performance of FAANG stocks before and after elections in 2016, 2020, and 2024 provides valuable insights into how market participants react to political outcomes and how these reactions shape price trajectories in the post-election period.
 
@@
 The November and December models share key hyperparameters optimized to balance complexity and generalizability:
+
 - **Max depth**: Set to 6, allowing the model to capture significant interactions without overfitting.
@@
 - **Boosting rounds**: Fixed at 100 to provide sufficient iterations for model convergence.
+
 XGBoost's built-in regularization techniques mitigate overfitting, especially important given the high-dimensional engineered features and the presence of missing values due to holidays and weekends.
@@
 #| eval: true
-#| label: fig-dec-error-distribution
+#| label: fig-dec-error-distrib
 #| fig-cap: "Histogram of residuals for the December model. Residuals are tightly centered around zero, indicating consistent predictions but limited sensitivity to abrupt stock price changes."
@@
 
-## Nature of Observational Data
+## Current Analysis: Surveys, Sampling, and Observational Data
+
+### Nature of Observational Data
 
@@
 
-## Sampling Frame and Representativeness
+### Sampling Frame and Representativeness
+
 The sampling frame encompasses all trading days from January 2020 to November 2024, providing a comprehensive record of FAANG stock activity during a period of significant market and political events.
@@
 
-### Strengths and Weaknesses
-
-A key strength of the dataset is its high-frequency nature, which allows the analysis to capture both short-term fluctuations and long-term trends.
+#### Strengths and Weaknesses
+
+A key strength of the data is its high-frequency nature, which allows the analysis to capture both short-term fluctuations and long-term trends.
 FAANG stocks, as market leaders, serve as valuable indicators of overall economic conditions and investor sentiment.
@@
 
-## Integrating Surveys with Observational Data
-
-Surveys provide forward-looking indicators that enhance the interpretation of stock price trends.
-For example, during significant events like the 2024 U.S.
-Presidential Election, surveys could capture investor sentiment regarding anticipated market impacts, which may not yet be reflected in observed price movements.
-
-### Proposed Survey Design
-
-To complement the observational stock data, investor sentiment surveys could include questions such as:
-
-- **Market Confidence**: “Do you expect FAANG stock prices to increase, decrease, or remain stable over the next month?”
-- **Election Impact**: “How has the 2024 U.S. Presidential Election influenced your investment decisions?”
-- **Risk Appetite**: “What is your likelihood of investing in technology stocks in the current market climate?”
-
-### Methodological Considerations:
-
-- **Sampling and Representativeness**:  Stratify respondents by demographics, investment experience, and portfolio size to ensure diverse perspectives.
-- **Weighting**: Apply post-survey weighting to align survey demographics with actual investor populations, similar to post-stratification in traditional surveys.
-
-### Strengths and Limitations
-
-Advantages of Survey Integration:
-
-- **Contextualization**: Surveys can help interpret stock price trends, particularly around major events like elections.
-- **Predictive Power**: Sentiment data may improve model accuracy by capturing market expectations and behavioral drivers.
-
-Limitations of Survey Integration:
-
-- **Response Bias**: Survey participants may exaggerate optimism or pessimism, especially during volatile periods.
-- **Sampling Complexity**: Ensuring representativeness across demographics, such as institutional vs. retail investors, requires significant resources and methodological rigor
-
-## Parallels Between Observational Data and Surveys
+### Parallels Between Observational Data and Surveys
 
@@
 
-## Addressing Data Challenges
-
-- **Temporal Dependence**: Stock prices are influenced by past trends.
-To mitigate this, features such as lagged prices (Lag_1) and rolling averages (Rolling_Mean_7) were used to capture momentum and smoothed trends effectively.
-However, these features may miss sudden shifts caused by external shocks or policy changes.
-
-- **Imputation for Missing Data**: Missing December 2024 values were filled through interpolation to ensure continuity.
-While necessary for analysis, this approach may underestimate volatility or fail to account for post-election market adjustments.
-Advanced time-series models could improve the accuracy of such imputations.
-
-- **Observational Bias**: Stock data is inherently influenced by external confounders, such as Federal Reserve policies, geopolitical tensions, or major corporate announcements.
-These confounders obscure causal relationships.
-Future models could include external predictors to better account for such influences, improving robustness.
-
-## Recommendations for Future Work
-
-1.
-**Survey Integration**: Combine investor sentiment surveys with stock data to capture qualitative dimensions of market behavior.
-Surveys can reveal forward-looking insights, such as investor confidence, that improve interpretability and predictive power.
-2. **Expanding Scope**: Extend the analysis to include sectors beyond FAANG stocks to enhance generalizability and uncover trends across diverse industries.
-3. **Improved Imputation**: Use dynamic time-series models to address missing values more accurately and capture volatility during critical events like elections or earnings announcements.
-
-By addressing these challenges and integrating surveys with observational data, future analyses can achieve a richer, more nuanced understanding of market dynamics.
-Such hybrid approach provides a comprehensive framework for interpreting stock price trends, making predictions more robust and actionable, especially during periods of significant economic or political change.
+## Idealized Methodology for FAANG Stock Prediction
+
+Building on the performance and limitations of our analysis, we now propose a robust framework for predicting FAANG stock prices by integrating observational data with investor sentiment surveys.
+The approach addresses limitations such as missing values, external confounders, and market volatility, incorporating best practices to ensure accurate, actionable predictions.
+It is designed to operate within a budget of $100,000.
+
+The link to our survey is as follows: "https://docs.google.com/forms/d/e/1FAIpQLScbhRrXNXA-70g6SplY8TcupSMQePFi3W6BbnwuAf1EQSrKJw/viewform?usp=sf_link"
+
+### Target Population, Sampling Frame, and Sample
+
+- **Target population**: Active retail investors who have traded FAANG stocks at least once in the past year.
+This group represents individuals most engaged with FAANG stock dynamics and whose sentiment provides actionable insights for predictive modeling.
+
+- **Sampling frame**: A comprehensive list of active retail investors derived from brokerage datasets.
+The frame includes demographic details (e.g., age, income, gender), trading behavior (e.g., frequency of trades), and portfolio composition (e.g., FAANG stock exposure).
+
+- **Sample**: A stratified sample of 1,000 respondents is selected to ensure proportional representation across key dimensions, including age (e.g., 18–34, 35–54, 55+), income level, portfolio size, and trading frequency.
+This stratification ensures diversity and reliability in capturing investor sentiment.
+
+### Sampling Methodology & Justification
+
+The sampling methodology employs stratified random sampling to ensure representation across diverse investor demographics. Respondents are stratified along the following dimensions:
+
+1. **Demographics**: Age, gender, and income levels.  
+2. **Portfolio Characteristics**: Size of portfolios and proportion allocated to FAANG stocks.  
+3. **Trading Behavior**: Frequency of trades, distinguishing between high-frequency and occasional investors.
+
+Stratification ensures the sample mirrors the diversity within the broader investor population.
+This method enhances generalizability while maintaining cost efficiency.
+With a sample size of 1,000, the methodology balances statistical rigor with practical feasibility.
+
+### Survey Design
+
+The survey is structured to capture critical insights into investor sentiment, preferences, and behaviors influencing FAANG stock price movements.
+It focuses on key dimensions such as market confidence, risk appetite, and reactions to external events like the 2024 U.S.
+Presidential Election.
+Designed with clarity and brevity, the survey ensures high-quality responses by minimizing respondent fatigue through concise, well-phrased questions.
+It incorporates a mix of quantitative and qualitative metrics, allowing respondents to express the intensity of their opinions on a scale-based format, providing a nuanced understanding of market sentiment.
+
+Our survey design also accounts for inclusivity and accessibility by being optimized for desktop and mobile platforms.
+To ensure reliability, validation questions are included to identify inattentive responses, and demographic questions enable stratified analysis across investor types.
+By framing questions within realistic contexts, such as recent market events and hypothetical FAANG performance scenarios, the survey gathers actionable data that enhances the interpretability and predictive power of stock price models.
+This design bridges the gap between quantitative trends and investor psychology, offering a comprehensive tool for integrating sentiment into financial forecasting.
+
+#### Cost and Feasibility
+
+The methodology is designed to operate within a $100,000 budget:
+
+- **Survey Development and Distribution**: $50,000 for platform costs and participant incentives.
+- **Data Acquisition**: $30,000 for accessing brokerage datasets and sampling frames.  
+- **Analysis and Reporting**: $20,000 for data integration, modeling, and reporting.
+
+### Data Validation
+
+To ensure reliable and representative survey data, we employ the following:
+
+- **Weighting**: Adjust responses to reflect population proportions using the formula:  
+$$
+  \text{Weight} = \frac{\text{Population Proportion}}{\text{Sample Proportion}}
+$$
+  This corrects for over- or under-representation within the sample.  
+- **Validation Questions**: Include items like “Select ‘Agree’ from the options below” to identify inattentive responses.  
+- **Unique IDs**: Assign each respondent a unique ID to prevent duplicates, retaining only the first response.  
+- **Real-Time Monitoring**: Track responses in real time to identify potential inconsistencies or fraudulent submissions.
+
+### Integration with Observational Data
+
+Survey responses will be integrated with historical stock price data by mapping survey-derived sentiment metrics, such as market confidence and risk appetite, to relevant time periods in the stock price dataset.
+For example, survey responses collected during November and December 2024 will be linked to corresponding stock price data, allowing the models to incorporate real-time investor sentiment alongside historical price trends.
+These sentiment metrics will be treated as additional predictive features in the modeling framework, enabling the machine learning models to capture the relationship between investor expectations and subsequent stock price movements.
+For instance, metrics like market confidence will be aligned with price trends to evaluate whether periods of high investor optimism correlate with upward price movements.
+Similarly, risk appetite data will help contextualize market volatility and identify patterns during periods of uncertainty.
+
+By combining survey insights with lagged prices, rolling averages, and volatility measures from the historical data, the models will leverage both forward-looking sentiment and past market behavior to make more robust and actionable predictions.
+This integration bridges the gap between investor psychology and quantitative market analysis, enhancing the models' ability to forecast stock price dynamics during volatile or event-driven periods.
+
+### Survey Demo
+
+Full survey can be found at "https://docs.google.com/forms/d/e/1FAIpQLScbhRrXNXA-70g6SplY8TcupSMQePFi3W6BbnwuAf1EQSrKJw/viewform?usp=sf_link" and under `FAANG-Stock-Forecast/other/survey`
+
+**Title: "Investor Sentiment Survey for FAANG Stock Predictions"**
+
+#### Survey Introduction
+We are conducting an important survey to understand investor opinions and predict FAANG stock price movements.
+Your participation will provide valuable insights into market sentiment and help improve stock price forecasting models.
+
+This survey will take about 5–10 minutes to complete.
+Your responses will remain completely anonymous and confidential.
+We are committed to protecting your privacy, and no personal information will be shared or used for any purpose beyond this research.
+
+#### Survey Questions:
+1. **Market Confidence**: “Do you expect FAANG stock prices to increase, decrease, or remain stable over the next month?”  
+   - Increase significantly  
+   - Increase slightly  
+   - Remain stable  
+   - Decrease slightly  
+   - Decrease significantly
+
+2. **Risk Appetite**: “What is your current risk appetite for investing in technology stocks?”  
+   - High  
+   - Moderate  
+   - Low
+
+3. **External Factors**: “To what extent have recent market events (e.g., 2024 U.S. Presidential Election, Federal Reserve decisions) influenced your investment decisions?”  
+   - A great deal  
+   - Somewhat  
+   - Not at all
+
+4. **Sector Preference**: “Besides FAANG stocks, which sectors are you considering for your portfolio over the next quarter?”  
+   - Technology (non-FAANG)  
+   - Healthcare  
+   - Energy  
+   - Consumer Goods  
+   - Other (please specify)
+
+5. **Economic Outlook**: “How do you perceive the broader economic climate for technology investments?”  
+   - Strong growth  
+   - Moderate growth  
+   - Stagnation  
+   - Decline
+
+Thank you for your time and contribution.
 
